### 2015-05-05 paper.tex
Added paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\textsc{18.409 - Algorithmic Aspects of Machine Learning} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Final Project \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Mike Salvato and Andrew Spielberg} % Your name

\date{\normalsize\today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

%----------------------------------------------------------------------------------------
%	PROBLEM 1
%----------------------------------------------------------------------------------------

\section{Approximate Clustering Without the Approximation}

They key point to this paper is to highlight the fact that, when talking about clustering, most people think of it in terms of minimizing some \emph{objective function} - e.g. minimizing the sum distance to nearby exemplar points (centroids) as in $k$-median clustering.  However, the actual \emph{hope} is that the output labelings produced by such a metric will be in some way "correct" - that is, that points are classified as some \emph{target} clustering up to some permutations of label.  More formally, this means that the hope is that if we find some clustering with an objective within a factor $c$ of the optimal, that we are then \emph{guaranteed} to have some clustering that is bounded by $\epsilon$ in a measure of "incorrectly clustered" points (we will define that measure later).  When we can assume this, we will see that we can derive efficient algorithms and guarantees for finding such $c$-approximation (and thus $\epsilon$ close in classification) clusterings.

\subsection{Introduction}

Spend one to two paragraph summarizing the main contributions and outline of this paper.

TODO: where do approximation algorithms come in?  What are they?  Talk about that here.

\subsection{Definitions}
As in the paper, we'll begin by defining a few important definitions:

First, let $C$ be a $k$-clustering, a partition of the set of $n$ points $S$ into a a set of subsets $\{C_1, C_2, ...C_k\}$, and let $C_T$ be the target clustering that we wish to find, and $d$ be the distance function between points on our metric space.  Let $\Phi$ be the objective function with minimum value $OPT_{\Phi}$, and let the distance between two clusterings $dist(C, C')$ be the 
fraction of points for which those two clusterings disagree on labels under the minimizing bijection between cluster labels.  See picture for an example (TODO: include pic).  Two clusterings $C, C'$ are $\epsilon$ close if $dist(C, C') < \epsilon$.

Now, we formalize the earlier assumption:

$S$ satisfies the $(c, \epsilon)$-assumption if \emph{all} clusterings $C$ with $\Phi(C) \leq c~OPT$ are $\epsilon$-close to $C_T$.

A few useful properties:

\begin{itemize}

\item \emph{uniqueness} - This actually doesn't even require the $(c, \epsilon)$ assumption, but merely that two clusterings are $\epsilon$-close.  In this case, if each cluster is of size at least $2n\epsilon$< the minimizing bijection is unique. TODO: why?  Also, is this worth keeping in?

\item Let's imagine that the target clustering $C_T$ that we wanted wasn't even the optimal clustering (satisfying $\Phi = OPT$), but some other clustering.  Then they these two clusterings are $\epsilon$-close.

\item Since $dist$ is a metric, it satisfies the triangle inequality, and therefore any clustering that that is within $(c, \epsilon)$ wrt $C_T$ also satisfies $(c, 2\epsilon)$ of the optimal clustering. (TODO: explain out with math?)

\end{itemize}

Following this section begins specific analyses of how the $(c, \epsilon)$ assumption leads to efficient algorithms for clustering for three commonly used objective functions ($k$-median, $k$-mean, $k$-min-sum).  However, before that, it's important to understand three very important limitations and implications of this setting.

\begin{enumerate}
\item $(c, \epsilon)$ properties do not imply $(kc, k\epsilon)$ properties in general for any of the three objective functions we'll analyze here.  To see this, let's look at a counterexample for the $k$-median example (TODO: should I generalize this?).  Let $1 \leq c_1 < c_2 $, and $\epsilon, \alpha > 0$.  Let the target clustering consist of a single cluster with $n(1 - 2\alpha$ points at a single location, and the remaining $k-1$ clusters each with $2\frac{\alpha n}{k-1}$ all at distance $1$ (TODO: draw a picture).  Each point in $C_1$ is greater than $c_2 n$ away from any other cluster.  Note that this clustering has objective $2\alpha n - (k - 1)$, in other words, all but one point in each cluster but $C_1$ contributes $1$ to the objective ($C_1$ contributes nothing since each point lies on the median).  Then, consider an alternate $c_2$ solution which splits $C_1$ into two clusters and merges clusters $C_2$ and $C_3$.  Since $n(\frac{1}{2} - \alpha)$ points from the first cluster are then misclassified (as it was split in two), clearly this has approximation error at least 
$\frac{1}{n}$ of that, and in fact, strictly greater.  If, in our original setup, we set the distance between points \emph{between} any two clusters to be $D > 1$ (which is defined to give this latter clustering a $c_2 OPT$ objective value).  But then, consider that any $c_1$ approximation cannot merge clusters (TODO: note that c2 was defined as the minimum that has a merging of clusters, can be seen by symmetry)).  error on fewer than $\frac{2\alpha n}{k-1} < \epsilon n$ (TODO: why?). Therefore, this is less than $k$ when we set $\alpha$ (TODO: explain that this defines the family) such that $k > 1 + \frac{2\alpha}{\epsilon} $.

TODO: draw this whole proof.

\item $c$-approximations under the $(c, \epsilon)$ assumption for $c > 1$ are as hard as finding a $c$-approximation in general. TODO: proof

\item Finally, finding a $c$-approximation under the $(1, \epsilon)$ assumption is NP-Hard. TODO: proof

\end{enumerate}

\section{Stability yields a PTAS for k-Median and k-Means Clustering}

This paper looks at a few statistics about how points are distributed and how it relates to optimal clusterings.  In particular, it studies notions of "stability," which describes bounds on how much adding or removing clusters from a clustering solution changes the objective function's value, and separability, which describes how far away clusters of points are from one another.  By characterizing point sets in terms of these statistics, the authors derive an algorithm for efficiently clustering the points and derive guarantees on the cluster quality, as well as prove its correctness.

\subsection{Introduction}
TODO: summarize the main contributions of this paper.

\subsection{Definitions}
The paper begins by describing a few statistics on the point distributions which are necessary for the subsequent analysis.  Let $OPT_i$ be the contribution to the objective function of the optimal clustering by cluster $i$, and let $OPT^{i \rightarrow j}$ denote the the cost when cluster $i$ is removed from the $OPT$ solution (TODO: unique?) and its pointed are then given to cluster $j$.  Intuitively, $OPT^{i \rightarrow j}$ corresponds to a (bad) solution of a $k-1$ clustering, and obviously $OPT(k)^{i \rightarrow j} \geq OPT(k-1)$, where $OPT(f)$ is the optimal solution of the $f$ cluster problem, and that $\geq$ sign will almost always be a $>$ sign in practice.  Let a clustering be $(1 + \alpha)$ weak-deletion stable if, $\forall i \neq j, OPT^{i \rightarrow j} > (1 + \alpha)OPT$ - in other words, removing any cluster and remapping those points to another single cluster severely contributes to the objective function, suggesting that that each cluster is in some sense \emph{important}.  We define

\end{document}

### 2015-05-05 paper.tex
Wrote up definitions for second paper

@@
 Let a clustering be $(1 + \alpha)$ weak-deletion stable if, $\forall i \neq j, OPT^{i \rightarrow j} > (1 + \alpha)OPT$ - in other words, removing any cluster and remapping those points to another single cluster severely contributes to the objective function, suggesting that that each cluster is in some sense \emph{important}.
-We define
+We define an optimal clustering solution to be $\beta$-distributed for $\beta > 0$ if for any $i$ and any point $x \notin C^i, d(x, c_i) \geq \beta \frac{OPT}{|C_i|}$ ; intuitively, a point outside a given cluster's distance to that cluster's center is sufficiently larger (measured by $\beta$ than the average distance of points within that cluster to its center.
+Specifically for this paper, we will focus on the $k$-median case, but analogous logic applies to the $k$-means case as well.
+
+Now, we prove that any $(1+\alpha)$-weakly deletion-stable $k$-median instance is $\frac{\alpha}{2}$-distributed,
+
+% and that any $(1+\alpha)$-weakly deletion-stable $k$-mean instance is $\frac{\alpha}{4}$-distributed.
+
+In each case, the proof will examine the difference $OPT^{i \rightarrow j} - OPT$.
+Each point $x \in C_i$ contributes $d(x, c_i)$ to $\Phi$, and on deleting $C_i$ and remapping its points to $C_j$, contributes $d(x, c_j)$.
+By the triangle inequality, $d(x, c_j) \leq d(x, c_i) + d(c_i, c_j)$.
+$d(x, c_j)$ is the contribution of the points that were in $C_i$ to $C_j$ upon deletion.
+Thus, the difference in contribution to $\Phi$ is the difference in cost upon swapping the points between the two clusters, which is $ d(x, c_j)  - d(x, c_i) \leq d(c_i, c_j)$.
+Summing up over the original cluster $C_i$,
+
+\begin{align}
+\sum_{k \in C_i} (d(x, c_j) )  - OPT_i \leq |C_i| d(c_i, c_j)
+\end{align}
+
+The increase (LHS) is also $ \geq \alpha OPT$ by $(1+\alpha)$-weakly deletion-stability, meaning $\alpha OPT \leq |C_i| d(c_i, c_j)$, or $d(c_i, c_j) \geq \frac{\alpha OPT}{|C_i|}$.
+Then, for any point $p \notin C_i, d(c_i, c_j) \leq d(c_i, p) + d(c_j, p)$, and this is $\leq 2d(c_i, p)$ (TODO: why?).
+Thus, applying this to the previous inequality, we get that $d(c_i, p) > \frac{\alpha OPT}{2 |C_i|}$; i.e.
+$\frac{alpha}{2}$ distributed.
+
+\subsection{Algorithm}
+TODO:
 

### 2015-05-09 paper.tex
some intro stuff

@@
 \usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
+
+%\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
 \usepackage[english]{babel} % English language/hyphenation
@@
 
+\usepackage{mathrsfs}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{bbm}
+\usepackage{amssymb} 
+\usepackage[mathscr]{euscript}
+\usepackage{parskip}
+
+\renewcommand{\vec}[1]{\mathbf{#1}}
+\newcommand{\vecsf}[1]{\textsf{\textbf{#1}}}
+
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
+\section{Introduction}
+
+Here we review work done on approximations for clustering algorithms, focusing on the work of \cite{firstpaper} and \cite{secondpaper}.
+The essential goal is that various data points about objects of interest, we can group those data points by which object they belong to.
+For example we may have a collection of face images, collect relative low dimensional data about the images, then cluster this data to try to determine images came from the same people by clustering them.
+However the problem of clustering points given a distance metric in general NP-hard.
+Additionally analysis of clustering alone does not take into account the relationship between the clusters and the underlying object they belong to.
+For example even if the images were clustered optimally, that clustering may not correctly group the images by which person's face is in the image.
+The work here will focus on approximating the underlying object a data point belongs to, using clustering methods.
+
+\subsection{Background}
+
+The algorithm we will focus on is the k-medians algorithm.
+The k-medians algorithm assumes a space $(X, d)$ for which $X$ is the set of possible points and distance $d : ({ X \choose 2}) \rightarrow R_{\geq 0}$ which satisfies the triangle inequality.
+We are then given a set of points $S \subset X$.
+The goal is to label the points in $S$ each into one of k clusters $C_1, C_2, ..., C_k$ with corresponding representative points $c_1, c_2, ..., c_k \in X$.
+This defines a clustering $\mathcal{C}$.
+The optimal objective clustering is that which minimizes $\Phi(\mathcal{C}) = \sum\limits_{i = 1}^k \sum_{x \in C_i} d(x, c_i)$.
+
+Thus far we have only addressed clustering based on the data points, but not the underlying objectives interest.
+The \textbf{objective} function $\Phi(\mathcal{C})$ is the cost of a clustering based only on the distance between points.
+However the optimal \textbf{target} clustering is that which partitions the points best based on the underlying objects the data points are drawn from.
+In the idealproblem statement the optimal objective clustering $\mathcal{C}^*$ and the optimal target clustering $\mathcal{C}_T$ are the same.
+However in practice this is not usually the case.
+In order to be able to perform analysis on the quality of a clustering \cite{firstpaper} define the $(c, \epsilon)$-property.
+If a clustering satisfies the $(c, \epsilon)$-property, then for all clusterings where $\Phi(\mathcal{C}) \leq c*\Phi(\mathcal{C}^*)$, $\mathcal{C}$ is also $\epsilon$-close to $\mathcal{C}_T$.
+By analyzing problems instances where this holds true for some values of $c$ and $\epsilon$, we can generate approximate clustering schema to achieve approximation solutions to discovering the target clustering.
+
+\subsection{Goals}
 

### 2015-05-10 paper.tex
introductions, clean up of content for paper two

@@
 \usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template
+\usepackage[margin=1.25in]{geometry}
 
@@
 
+\subsection{k-medians}
 The algorithm we will focus on is the k-medians algorithm.
@@
 
+Let the distance between two clusterings $dist(C, C')$ be the 
+fraction of points for which those two clusterings disagree on labels under the minimizing bijection between cluster labels, and let $\text{OPT}_\Phi = \Phi(\mathcal{C}^*)$
+
+\subsection{($c, \delta$)-property}
 Thus far we have only addressed clustering based on the data points, but not the underlying objectives interest.
@@
 However the optimal \textbf{target} clustering is that which partitions the points best based on the underlying objects the data points are drawn from.
-In the idealproblem statement the optimal objective clustering $\mathcal{C}^*$ and the optimal target clustering $\mathcal{C}_T$ are the same.
+In the ideal problem statement the optimal objective clustering $\mathcal{C}^*$ and the optimal target clustering $\mathcal{C}_T$ are the same.
 However in practice this is not usually the case.
-In order to be able to perform analysis on the quality of a clustering \cite{firstpaper} define the $(c, \epsilon)$-property.
-If a clustering satisfies the $(c, \epsilon)$-property, then for all clusterings where $\Phi(\mathcal{C}) \leq c*\Phi(\mathcal{C}^*)$, $\mathcal{C}$ is also $\epsilon$-close to $\mathcal{C}_T$.
-By analyzing problems instances where this holds true for some values of $c$ and $\epsilon$, we can generate approximate clustering schema to achieve approximation solutions to discovering the target clustering.
+In order to be able to perform analysis on the quality of a clustering \cite{firstpaper} define the $(c, \delta)$-property.
+If a clustering satisfies the $(c, \delta)$-property, then for all clusterings where $\Phi(\mathcal{C}) \leq c*\text{OPT}_\Phi$, $\mathcal{C}$ is also $\delta$-close to $\mathcal{C}_T$.
+By analyzing problems instances where this holds true for some values of $c$ and $\delta$, we can generate approximate clustering schema to achieve approximation solutions to discovering the target clustering.
 
@@
 
-\section{Approximate Clustering Without the Approximation}
-
-They key point to this paper is to highlight the fact that, when talking about clustering, most people think of it in terms of minimizing some \emph{objective function} - e.g.
-minimizing the sum distance to nearby exemplar points (centroids) as in $k$-median clustering.
-However, the actual \emph{hope} is that the output labelings produced by such a metric will be in some way "correct" - that is, that points are classified as some \emph{target} clustering up to some permutations of label.
-More formally, this means that the hope is that if we find some clustering with an objective within a factor $c$ of the optimal, that we are then \emph{guaranteed} to have some clustering that is bounded by $\epsilon$ in a measure of "incorrectly clustered" points (we will define that measure later).
-When we can assume this, we will see that we can derive efficient algorithms and guarantees for finding such $c$-approximation (and thus $\epsilon$ close in classification) clusterings.
-
-\subsection{Introduction}
-
-Spend one to two paragraph summarizing the main contributions and outline of this paper.
-
-TODO: where do approximation algorithms come in?  What are they?  Talk about that here.
-
-\subsection{Definitions}
-As in the paper, we'll begin by defining a few important definitions:
-
-First, let $C$ be a $k$-clustering, a partition of the set of $n$ points $S$ into a a set of subsets $\{C_1, C_2, ...C_k\}$, and let $C_T$ be the target clustering that we wish to find, and $d$ be the distance function between points on our metric space.
-Let $\Phi$ be the objective function with minimum value $OPT_{\Phi}$, and let the distance between two clusterings $dist(C, C')$ be the
-fraction of points for which those two clusterings disagree on labels under the minimizing bijection between cluster labels.
-See picture for an example (TODO: include pic).
-Two clusterings $C, C'$ are $\epsilon$ close if $dist(C, C') < \epsilon$.
-
-Now, we formalize the earlier assumption:
-
-$S$ satisfies the $(c, \epsilon)$-assumption if \emph{all} clusterings $C$ with $\Phi(C) \leq c~OPT$ are $\epsilon$-close to $C_T$.
-
-A few useful properties:
+The goal of this work is to develop a minimal runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
+It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{thethingthepapercitesJMS02}.
+In this review we will show explore the following results:
 
@@
 
-\item \emph{uniqueness} - This actually doesn't even require the $(c, \epsilon)$ assumption, but merely that two clusterings are $\epsilon$-close.
-In this case, if each cluster is of size at least $2n\epsilon$< the minimizing bijection is unique.
+\item Assume we have a problem with a $(1+\alpha, \delta)$-property, and that all target clusters have size greater than $\delta n$. There exists a PTAS $\delta$-close to the target.
+
+\item Assume that the (k-1)-medians optimal cluster is more expensive than the k optimal clustering by a factor of $(1+\alpha)$ where $\alpha > 0$.
+Then it is possible to achieve a ($1+\epsilon$) approximation to the optimal k-medians in time polynomial in $n$ and $k$, and exponential in $1/\epsilon$ and $1/\alpha$, where $n = |S|$.
+
+\end{itemize}
+%
+%\section{Approximate Clustering Without the Approximation}
+%
+%
+%They key point to this paper is to highlight the fact that, when talking about clustering, most people think of it in terms of minimizing some \emph{objective function} - e.g.
+minimizing the sum distance to nearby exemplar points (centroids) as in $k$-median clustering.
+However, the actual \emph{hope} is that the output labelings produced by such a metric will be in some way "correct" - that is, that points are classified as some \emph{target} clustering up to some permutations of label.
+More formally, this means that the hope is that if we find some clustering with an objective within a factor $c$ of the optimal, that we are then \emph{guaranteed} to have some clustering that is bounded by $\delta$ in a measure of "incorrectly clustered" points (we will define that measure later).
+When we can assume this, we will see that we can derive efficient algorithms and guarantees for finding such $c$-approximation (and thus $\delta$ close in classification) clusterings.
+%
+%\subsection{Introduction}
+%
+%Spend one to two paragraph summarizing the main contributions and outline of this paper.
+%
+%TODO: where do approximation algorithms come in?  What are they?  Talk about that here.
+%
+%\subsection{Definitions}
+%As in the paper, we'll begin by defining a few important definitions:
+%
+%First, let $C$ be a $k$-clustering, a partition of the set of $n$ points $S$ into a a set of subsets $\{C_1, C_2, ...C_k\}$, and let $C_T$ be the target clustering that we wish to find, and $d$ be the distance function between points on our metric space.
+Let $\Phi$ be the objective function with minimum value $OPT_{\Phi}$, and let the distance between two clusterings $dist(C, C')$ be the
+%fraction of points for which those two clusterings disagree on labels under the minimizing bijection between cluster labels.
+See picture for an example (TODO: include pic).
+Two clusterings $C, C'$ are $\delta$ close if $dist(C, C') < \delta$.
+%
+%Now, we formalize the earlier assumption:
+%
+%$S$ satisfies the $(c, \delta)$-assumption if \emph{all} clusterings $C$ with $\Phi(C) \leq c~OPT$ are $\delta$-close to $C_T$.
+%
+%A few useful properties:
+%
+%\begin{itemize}
+%
+%\item \emph{uniqueness} - This actually doesn't even require the $(c, \delta)$ assumption, but merely that two clusterings are $\delta$-close.
+In this case, if each cluster is of size at least $2n\delta$< the minimizing bijection is unique.
 TODO: why?
 Also, is this worth keeping in?
-
-\item Let's imagine that the target clustering $C_T$ that we wanted wasn't even the optimal clustering (satisfying $\Phi = OPT$), but some other clustering.
-Then they these two clusterings are $\epsilon$-close.
-
-\item Since $dist$ is a metric, it satisfies the triangle inequality, and therefore any clustering that that is within $(c, \epsilon)$ wrt $C_T$ also satisfies $(c, 2\epsilon)$ of the optimal clustering.
+%
+%\item Let's imagine that the target clustering $C_T$ that we wanted wasn't even the optimal clustering (satisfying $\Phi = OPT$), but some other clustering.
+Then they these two clusterings are $\delta$-close.
+%
+%\item Since $dist$ is a metric, it satisfies the triangle inequality, and therefore any clustering that that is within $(c, \delta)$ wrt $C_T$ also satisfies $(c, 2\delta)$ of the optimal clustering.
 (TODO: explain out with math?)
-
-\end{itemize}
-
-Following this section begins specific analyses of how the $(c, \epsilon)$ assumption leads to efficient algorithms for clustering for three commonly used objective functions ($k$-median, $k$-mean, $k$-min-sum).
+%
+%
+%
+%\end{itemize}
+%
+%Following this section begins specific analyses of how the $(c, \delta)$ assumption leads to efficient algorithms for clustering for three commonly used objective functions ($k$-median, $k$-mean, $k$-min-sum).
 However, before that, it's important to understand three very important limitations and implications of this setting.
-
-\begin{enumerate}
-\item $(c, \epsilon)$ properties do not imply $(kc, k\epsilon)$ properties in general for any of the three objective functions we'll analyze here.
+%
+%\begin{enumerate}
+%\item $(c, \delta)$ properties do not imply $(kc, k\delta)$ properties in general for any of the three objective functions we'll analyze here.
 To see this, let's look at a counterexample for the $k$-median example (TODO: should I generalize this?).
-Let $1 \leq c_1 < c_2 $, and $\epsilon, \alpha > 0$.
+Let $1 \leq c_1 < c_2 $, and $\delta, \alpha > 0$.
 Let the target clustering consist of a single cluster with $n(1 - 2\alpha$ points at a single location, and the remaining $k-1$ clusters each with $2\frac{\alpha n}{k-1}$ all at distance $1$ (TODO: draw a picture).
@@
 Since $n(\frac{1}{2} - \alpha)$ points from the first cluster are then misclassified (as it was split in two), clearly this has approximation error at least
-$\frac{1}{n}$ of that, and in fact, strictly greater.
+%$\frac{1}{n}$ of that, and in fact, strictly greater.
 If, in our original setup, we set the distance between points \emph{between} any two clusters to be $D > 1$ (which is defined to give this latter clustering a $c_2 OPT$ objective value).
 But then, consider that any $c_1$ approximation cannot merge clusters (TODO: note that c2 was defined as the minimum that has a merging of clusters, can be seen by symmetry)).
-error on fewer than $\frac{2\alpha n}{k-1} < \epsilon n$ (TODO: why?).
-Therefore, this is less than $k$ when we set $\alpha$ (TODO: explain that this defines the family) such that $k > 1 + \frac{2\alpha}{\epsilon} $.
-
-TODO: draw this whole proof.
-
-\item $c$-approximations under the $(c, \epsilon)$ assumption for $c > 1$ are as hard as finding a $c$-approximation in general. TODO: proof
-
-\item Finally, finding a $c$-approximation under the $(1, \epsilon)$ assumption is NP-Hard. TODO: proof
-
-\end{enumerate}
+error on fewer than $\frac{2\alpha n}{k-1} < \delta n$ (TODO: why?).
+Therefore, this is less than $k$ when we set $\alpha$ (TODO: explain that this defines the family) such that $k > 1 + \frac{2\alpha}{\delta} $.
+%
+%TODO: draw this whole proof.
+%
+%
+%
+%\item $c$-approximations under the $(c, \delta)$ assumption for $c > 1$ are as hard as finding a $c$-approximation in general. TODO: proof
+%
+%\item Finally, finding a $c$-approximation under the $(1, \delta)$ assumption is NP-Hard. TODO: proof
+%
+%\end{enumerate}
 
@@
 Let $OPT_i$ be the contribution to the objective function of the optimal clustering by cluster $i$, and let $OPT^{i \rightarrow j}$ denote the the cost when cluster $i$ is removed from the $OPT$ solution (TODO: unique?) and its pointed are then given to cluster $j$.
-Intuitively, $OPT^{i \rightarrow j}$ corresponds to a (bad) solution of a $k-1$ clustering, and obviously $OPT(k)^{i \rightarrow j} \geq OPT(k-1)$, where $OPT(f)$ is the optimal solution of the $f$ cluster problem, and that $\geq$ sign will almost always be a $>$ sign in practice.
-Let a clustering be $(1 + \alpha)$ weak-deletion stable if, $\forall i \neq j, OPT^{i \rightarrow j} > (1 + \alpha)OPT$ - in other words, removing any cluster and remapping those points to another single cluster severely contributes to the objective function, suggesting that that each cluster is in some sense \emph{important}.
-We define an optimal clustering solution to be $\beta$-distributed for $\beta > 0$ if for any $i$ and any point $x \notin C^i, d(x, c_i) \geq \beta \frac{OPT}{|C_i|}$ ; intuitively, a point outside a given cluster's distance to that cluster's center is sufficiently larger (measured by $\beta$ than the average distance of points within that cluster to its center.
+Intuitively, $OPT^{i \rightarrow j}$ corresponds to a (bad) solution of a $k-1$ clustering, and $OPT(k)^{i \rightarrow j} \geq OPT(k-1)$, where $OPT(f)$ is the optimal solution of the $f$ cluster problem.
+Let a clustering be $(1 + \alpha)$ weak-deletion stable if, $\forall i \neq j, OPT^{i \rightarrow j} > (1 + \alpha)OPT$ - in other words, removing any cluster and remapping those points to another single cluster severely contributes to the objective function, suggesting that that each cluster is significant to the solution.
+Intuitively if this is not the case, then one could reasonably use the ($k-1$) solution in place of the $k$ solution.
+
+  We define an optimal clustering solution $\mathcal{C}^*$ to be $\beta$-distributed for $\beta > 0$ if for any $i$ and any point $x \notin C^*_i, d(x, c^*_i) \geq \beta \frac{OPT}{|C^*_i|}$.
+Intuitively, this provides a factor $\beta$ for the distance that each point is from another cluster's center, relative to the average distance a point is from its own cluster center.
 Specifically for this paper, we will focus on the $k$-median case, but analogous logic applies to the $k$-means case as well.
@@
 In each case, the proof will examine the difference $OPT^{i \rightarrow j} - OPT$.
-Each point $x \in C_i$ contributes $d(x, c_i)$ to $\Phi$, and on deleting $C_i$ and remapping its points to $C_j$, contributes $d(x, c_j)$.
-By the triangle inequality, $d(x, c_j) \leq d(x, c_i) + d(c_i, c_j)$.
-$d(x, c_j)$ is the contribution of the points that were in $C_i$ to $C_j$ upon deletion.
-Thus, the difference in contribution to $\Phi$ is the difference in cost upon swapping the points between the two clusters, which is $ d(x, c_j)  - d(x, c_i) \leq d(c_i, c_j)$.
-Summing up over the original cluster $C_i$,
+Each point $x \in C^*_i$ contributes $d(x, c^*_i)$ to $\Phi$, and on deleting $C^*_i$ and remapping its points to $C^*_j$, contributes $d(x, c^*_j)$.
+By the triangle inequality, $d(x, c^*_j) \leq d(x, c^*_i) + d(c^*_i, c^*_j)$.
+Thus, the difference in contribution to $\Phi$ is the difference in cost upon swapping the points between the two clusters, which is $ d(x, c^*_j)  - d(x, c^*_i) \leq d(c^*_i, c^*_j)$.
+Summing up over the original cluster $C^*_i$,
 
 \begin{align}
-\sum_{k \in C_i} (d(x, c_j) )  - OPT_i \leq |C_i| d(c_i, c_j)
+\sum_{x \in C^*_i} (d(x, c^*_j) )  - OPT_i \leq |C^*_i| d(c^*_i, c^*_j)
 \end{align}
 
-The increase (LHS) is also $ \geq \alpha OPT$ by $(1+\alpha)$-weakly deletion-stability, meaning $\alpha OPT \leq |C_i| d(c_i, c_j)$, or $d(c_i, c_j) \geq \frac{\alpha OPT}{|C_i|}$.
-Then, for any point $p \notin C_i, d(c_i, c_j) \leq d(c_i, p) + d(c_j, p)$, and this is $\leq 2d(c_i, p)$ (TODO: why?).
-Thus, applying this to the previous inequality, we get that $d(c_i, p) > \frac{\alpha OPT}{2 |C_i|}$; i.e.
+The quantity on the LHS represents the increase in cost, which we know is $ > \alpha OPT$ by $(1+\alpha)$-weakly deletion-stability.
+This means $\alpha OPT < |C^*_i| d(c^*_i, c^*_j)$, or $d(c^*_i, c^*_j) >\frac{\alpha OPT}{|C^*_i|}$.
+Then, for any point $p \in C^*_j, d(c^*_i, c^*_j) \leq d(c^*_i, p) + d(c^*_j, p) \leq 2d(c^*_i, p)$.
+The former inequality is due to the triangle inequality.
+Thus, applying this to the previous inequality, we get that $d(c^*_i, p) > \frac{\alpha OPT}{2 |C^*_i|}$; i.e.
 $\frac{alpha}{2}$ distributed.

### 2015-05-10 paper.tex
added some more definitions for second paper, move in properties of c delta

@@
 The goal is to label the points in $S$ each into one of k clusters $C_1, C_2, ..., C_k$ with corresponding representative points $c_1, c_2, ..., c_k \in X$.
-This defines a clustering $\mathcal{C}$.
-The optimal objective clustering is that which minimizes $\Phi(\mathcal{C}) = \sum\limits_{i = 1}^k \sum_{x \in C_i} d(x, c_i)$.
+This defines a clustering $C$.
+The optimal objective clustering is that which minimizes $\Phi(C) = \sum\limits_{i = 1}^k \sum_{x \in C_i} d(x, c_i)$.
 
 Let the distance between two clusterings $dist(C, C')$ be the 
-fraction of points for which those two clusterings disagree on labels under the minimizing bijection between cluster labels, and let $\text{OPT}_\Phi = \Phi(\mathcal{C}^*)$
+fraction of points for which those two clusterings disagree on labels under the minimizing bijection between cluster labels, and let $\text{OPT}_\Phi = \Phi(C^*)$
 
@@
 Thus far we have only addressed clustering based on the data points, but not the underlying objectives interest.
-The \textbf{objective} function $\Phi(\mathcal{C})$ is the cost of a clustering based only on the distance between points.
+The \textbf{objective} function $\Phi(C)$ is the cost of a clustering based only on the distance between points.
 However the optimal \textbf{target} clustering is that which partitions the points best based on the underlying objects the data points are drawn from.
-In the ideal problem statement the optimal objective clustering $\mathcal{C}^*$ and the optimal target clustering $\mathcal{C}_T$ are the same.
+In the ideal problem statement the optimal objective clustering $C^*$ and the optimal target clustering $C^T$ are the same.
 However in practice this is not usually the case.
-In order to be able to perform analysis on the quality of a clustering \cite{firstpaper} define the $(c, \delta)$-property.
-If a clustering satisfies the $(c, \delta)$-property, then for all clusterings where $\Phi(\mathcal{C}) \leq c*\text{OPT}_\Phi$, $\mathcal{C}$ is also $\delta$-close to $\mathcal{C}_T$.
+In order to be able to perform analysis on the quality of a clustering \cite{firstpaper} define the $(c, \delta)$-property (alternatively ($c, \delta$) stability)).
+If a clustering satisfies the $(c, \delta)$-property, then for all clusterings where $\Phi(C) \leq c*\text{OPT}_\Phi$, $C$ is also $\delta$-close to $C^T$.
 By analyzing problems instances where this holds true for some values of $c$ and $\delta$, we can generate approximate clustering schema to achieve approximation solutions to discovering the target clustering.
 
-\subsection{Goals}
-
-The goal of this work is to develop a minimal runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
-It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{thethingthepapercitesJMS02}.
-In this review we will show explore the following results:
+ DO WE NEED THESE? THE LATTER TWO ARE PRETTY MUCH HOW IT'S DEFINED... (I know it's in the paper, I'm not sure I really get why)
+ A few useful properties:
 
@@
 
-\item Assume we have a problem with a $(1+\alpha, \delta)$-property, and that all target clusters have size greater than $\delta n$. There exists a PTAS $\delta$-close to the target.
+\item Let's imagine that the target clustering $C_T$ isn't the optimal clustering (satisfying $\Phi = OPT$), but some other clustering.  Then they these two clusterings are $\delta$-close.
 
-\item Assume that the (k-1)-medians optimal cluster is more expensive than the k optimal clustering by a factor of $(1+\alpha)$ where $\alpha > 0$.
-Then it is possible to achieve a ($1+\epsilon$) approximation to the optimal k-medians in time polynomial in $n$ and $k$, and exponential in $1/\epsilon$ and $1/\alpha$, where $n = |S|$.
+\item Since $dist$ is a metric, it satisfies the triangle inequality, and therefore any clustering that that is within $(c, \delta)$ wrt $C_T$ also satisfies $(c, 2\delta)$ of the optimal clustering.
+%(TODO: explain out with math?) I don't think it's necessary.
 
 \end{itemize}
-%
-%\section{Approximate Clustering Without the Approximation}
-%
-%
-%They key point to this paper is to highlight the fact that, when talking about clustering, most people think of it in terms of minimizing some \emph{objective function} - e.g.
-minimizing the sum distance to nearby exemplar points (centroids) as in $k$-median clustering.
-However, the actual \emph{hope} is that the output labelings produced by such a metric will be in some way "correct" - that is, that points are classified as some \emph{target} clustering up to some permutations of label.
-More formally, this means that the hope is that if we find some clustering with an objective within a factor $c$ of the optimal, that we are then \emph{guaranteed} to have some clustering that is bounded by $\delta$ in a measure of "incorrectly clustered" points (we will define that measure later).
-When we can assume this, we will see that we can derive efficient algorithms and guarantees for finding such $c$-approximation (and thus $\delta$ close in classification) clusterings.
-%
-%\subsection{Introduction}
-%
-%Spend one to two paragraph summarizing the main contributions and outline of this paper.
-%
-%TODO: where do approximation algorithms come in?  What are they?  Talk about that here.
-%
-%\subsection{Definitions}
-%As in the paper, we'll begin by defining a few important definitions:
-%
-%First, let $C$ be a $k$-clustering, a partition of the set of $n$ points $S$ into a a set of subsets $\{C_1, C_2, ...C_k\}$, and let $C_T$ be the target clustering that we wish to find, and $d$ be the distance function between points on our metric space.
-Let $\Phi$ be the objective function with minimum value $OPT_{\Phi}$, and let the distance between two clusterings $dist(C, C')$ be the
-%fraction of points for which those two clusterings disagree on labels under the minimizing bijection between cluster labels.
-See picture for an example (TODO: include pic).
-Two clusterings $C, C'$ are $\delta$ close if $dist(C, C') < \delta$.
-%
-%Now, we formalize the earlier assumption:
-%
-%$S$ satisfies the $(c, \delta)$-assumption if \emph{all} clusterings $C$ with $\Phi(C) \leq c~OPT$ are $\delta$-close to $C_T$.
-%
-%A few useful properties:
-%
-%\begin{itemize}
-%
-%\item \emph{uniqueness} - This actually doesn't even require the $(c, \delta)$ assumption, but merely that two clusterings are $\delta$-close.
-In this case, if each cluster is of size at least $2n\delta$< the minimizing bijection is unique.
-TODO: why?
-Also, is this worth keeping in?
-%
-%\item Let's imagine that the target clustering $C_T$ that we wanted wasn't even the optimal clustering (satisfying $\Phi = OPT$), but some other clustering.
-Then they these two clusterings are $\delta$-close.
-%
-%\item Since $dist$ is a metric, it satisfies the triangle inequality, and therefore any clustering that that is within $(c, \delta)$ wrt $C_T$ also satisfies $(c, 2\delta)$ of the optimal clustering.
-(TODO: explain out with math?)
-%
-%
-%
-%\end{itemize}
-%
+
+% Do we need all these? They're basically the strawmen, so I'm not sure they're the best use of space.
+
 %Following this section begins specific analyses of how the $(c, \delta)$ assumption leads to efficient algorithms for clustering for three commonly used objective functions ($k$-median, $k$-mean, $k$-min-sum).
@@
 
+\subsection{Goals}
+
+The goal of this work is to develop a minimal runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
+It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{thethingthepapercitesJMS02}.
+In this review we will show explore the following results:
+
+\begin{itemize}
+
+\item Assume we have a problem with a $(1+\alpha, \delta)$-property, and that all target clusters have size greater than $\delta n$. There exists a PTAS $\delta$-close to the target.
+
+\item Assume that the (k-1)-medians optimal cluster is more expensive than the k optimal clustering by a factor of $(1+\alpha)$ where $\alpha > 0$.
+Then it is possible to achieve a ($1+\epsilon$) approximation to the optimal k-medians in time polynomial in $n$ and $k$, and exponential in $1/\epsilon$ and $1/\alpha$, where $n = |S|$.
+
+\end{itemize}
+
 \section{Stability yields a PTAS for k-Median and k-Means Clustering}
@@
 
-  We define an optimal clustering solution $\mathcal{C}^*$ to be $\beta$-distributed for $\beta > 0$ if for any $i$ and any point $x \notin C^*_i, d(x, c^*_i) \geq \beta \frac{OPT}{|C^*_i|}$.
+  We define an optimal clustering solution $C^*$ to be $\beta$-distributed for $\beta > 0$ if for any $i$ and any point $x \notin C^*_i, d(x, c^*_i) \geq \beta \frac{OPT}{|C^*_i|}$.
 Intuitively, this provides a factor $\beta$ for the distance that each point is from another cluster's center, relative to the average distance a point is from its own cluster center.
@@
 
+\subsubsection{Relationship to the $(c, \delta)$-property}
+
+We show that a k-median cluster with ($1+\alpha, \delta$) stability in which all clusters have size greater than $\delta n$ have $(1+\alpha)$ weak deletion-stability.
+
+We use $C^{(i\rightarrow j)}$ to indicate the clustering obtained by taking all the points in cluster $C_i^*$ and moving them to cluster $C_j^*$.
+Because every cluster has at least $\delta n$ points, and $C^{(i \rightarrow j)}$ has one fewer cluster than $C^T$ we know the distance from $C^T$ and $C^{(i\rightarrow j)}$ is greater than $\delta$.
+Therefor the cost of $C^{(i\rightarrow j)}$ is at least $(1 + \alpha)\text{OPT}$.
+
 \subsection{Algorithm}

### 2015-05-10 paper.tex
Added intuition for algorithm

@@
 Because every cluster has at least $\delta n$ points, and $C^{(i \rightarrow j)}$ has one fewer cluster than $C^T$ we know the distance from $C^T$ and $C^{(i\rightarrow j)}$ is greater than $\delta$.
-Therefor the cost of $C^{(i\rightarrow j)}$ is at least $(1 + \alpha)\text{OPT}$.
+Therefore the cost of $C^{(i\rightarrow j)}$ is at least $(1 + \alpha)\text{OPT}$.
 
 \subsection{Algorithm}
-TODO:
+First, like in the paper, we'll begin by briefly discussing the intuition of the algorithm.
+
+Let's begin by partitioning our set of clusters into two types.
+There are  "cheap" clusters and "expensive" clusters, with the idea being that the expensive clusters contribute a relatively much larger amount than the cheaper clusters; in other words, we have mostly densely packed groups of points, and a few not-so densely packed groups.
+The key idea is to assume (and later we prove that the assumption is valid) that the number of expensive clusters is \emph{bounded}.
+We then have a method for clustering the cheap point sets, assuming we can partition cheap and expensive points - guess the number of points in the cheap clusters, derive a relationship from the number of points to the radius $r$ of that point set, and then draw balls of radius $r$ around the points in each of those sets.
+In order to cluster the expensive points, we then use brute force to guess each of the remaining cluster centers, and since the number of remaining clusters is bounded and small, this brute force operation should not contribute significantly to the runtime.
+This all hinges on being able to partition the points into cheap and expensive sets, which arises naturally out of the procedure.
+In practice, the algorithm attempts to create a list of possible subsets of points (which could overlap in membership), and iteratively refine that list to create our $k$ clusters.
+
+TODO: verify all this.
+
+Because we assume the clustering is $\beta$-distributed,
 

### 2015-05-10 paper.tex
a word"

--- 2015-05-10 paper.tex
@@
 
+In order to begin our analysis of clustering, we first define what clustering means exactly, and the relationship it holds to solving an underlying data analysis problem.
+
 \subsection{k-medians}
@@
 However in practice this is not usually the case.
-In order to be able to perform analysis on the quality of a clustering \cite{firstpaper} define the $(c, \delta)$-property (alternatively ($c, \delta$) stability)).
+In order to be able to perform analysis on the quality of a clustering \cite{firstpaper} define the $(c, \delta)$-property.
 If a clustering satisfies the $(c, \delta)$-property, then for all clusterings where $\Phi(C) \leq c*\text{OPT}_\Phi$, $C$ is also $\delta$-close to $C^T$.
@@
 
- DO WE NEED THESE? THE LATTER TWO ARE PRETTY MUCH HOW IT'S DEFINED... (I know it's in the paper, I'm not sure I really get why)
- A few useful properties:
+ A couple of useful properties:
 
@@
 
-\item Let's imagine that the target clustering $C_T$ isn't the optimal clustering (satisfying $\Phi = OPT$), but some other clustering.  Then they these two clusterings are $\delta$-close.
+\item In the case where the target clustering $C^T$ isn't the optimal clustering $C^*$, these two clusterings are $\delta$-close.
 
@@
 
-\subsection{Introduction}
-TODO: summarize the main contributions of this paper.
-
 \subsection{Definitions}
@@
 Thus, applying this to the previous inequality, we get that $d(c^*_i, p) > \frac{\alpha OPT}{2 |C^*_i|}$; i.e.
-$\frac{alpha}{2}$ distributed.
+$\frac{\alpha}{2}$ distributed.
 

### 2015-05-10 paper.tex
Defining inner ring

--- 2015-05-10 paper.tex
@@
 \subsection{Algorithm}
-First, like in the paper, we'll begin by briefly discussing the intuition of the algorithm.
+First, like in the paper, we'll begin by briefly discussing the intuition of the $(1 + \epsilon)$ approximation algorithm.
 
@@
 
+\subsubsection{Inner Ring}
+
+Define a \emph{cheap cluster} as one where $OPT_i \leq \frac{\beta \epsilon OPT}{32}$, meaning it's less than $\frac{32}{\beta \epsilon}$ of $OPT$.
+Let the \emph{inner ring} of $C_i$ be the set of points $x$ such that $d(x, c_i) \leq \frac{ \beta OPT }{8 |C_i|}$.
+If $C_i$ is cheap, then no more than $\frac{\epsilon}{4}$ points reside outside its inner ring.
+By contradiction, assume that more than $\frac{\epsilon}{4}$ fraction of points exist outside of the inner ring.
+$OPT_i > (points outside inner ring) * (cost of points)$, and the first term in that product is lower bounded by $\frac{\epsilon |C_i|}{4}$ and that second term is lower bounded by $\frac{\beta OPT}{8 |C_i|}$.
+Thus, $OPT_i > \frac{\beta \epsilon OPT }{32}$.
+But all cheap clusters have cost less than $\frac{\beta \epsilon OPT}{32}$, so we have a contradiction, and thus cheap clusters only have at most $\frac{\epsilon}{4}$ points outside their inner rings.
+
+TODO: what on Earth does Markov's inequality have to do with this?  Is this a typo in the paper?
+
+\subsubsection{The Algorithm}
+
+Blah blah blah describe algorithm here
+
+\subsubsection{Bounds on Components of Cheap Clusters}
+We now move into the proof of correctness, and start by proving a useful lemma here.
+As mentioned before, our goal is to find lists of \emph{components}, or subsets of points, from which we hope we can derive our clusters.
+We now show that, if we bound
+
 \end{document}

### 2015-05-10 paper.tex
Algorithm summarization

@@
 
-\subsection{Algorithm}
-First, like in the paper, we'll begin by briefly discussing the intuition of the $(1 + \epsilon)$ approximation algorithm.
-
-Let's begin by partitioning our set of clusters into two types.
-There are  "cheap" clusters and "expensive" clusters, with the idea being that the expensive clusters contribute a relatively much larger amount than the cheaper clusters; in other words, we have mostly densely packed groups of points, and a few not-so densely packed groups.
-The key idea is to assume (and later we prove that the assumption is valid) that the number of expensive clusters is \emph{bounded}.
-We then have a method for clustering the cheap point sets, assuming we can partition cheap and expensive points - guess the number of points in the cheap clusters, derive a relationship from the number of points to the radius $r$ of that point set, and then draw balls of radius $r$ around the points in each of those sets.
-In order to cluster the expensive points, we then use brute force to guess each of the remaining cluster centers, and since the number of remaining clusters is bounded and small, this brute force operation should not contribute significantly to the runtime.
-This all hinges on being able to partition the points into cheap and expensive sets, which arises naturally out of the procedure.
-In practice, the algorithm attempts to create a list of possible subsets of points (which could overlap in membership), and iteratively refine that list to create our $k$ clusters.
-
-TODO: verify all this.
-
-Because we assume the clustering is $\beta$-distributed,
-
 \subsubsection{Inner Ring}
@@
 
-\subsubsection{The Algorithm}
+\subsection{Algorithm}
+First, like in the paper, we'll begin by briefly discussing the intuition of the $(1 + \epsilon)$ approximation algorithm.
 
-Blah blah blah describe algorithm here
+Let's begin by partitioning our set of clusters into two types.
+There are  "cheap" clusters and "expensive" clusters, with the idea being that the expensive clusters contribute a relatively much larger amount than the cheaper clusters; in other words, we have mostly densely packed groups of points, and a few not-so densely packed groups.
+The key idea is to assume (and later we prove that the assumption is valid) that the number of expensive clusters is \emph{bounded}.
+We then have a method for clustering the cheap point sets, assuming we can partition cheap and expensive points - guess the number of points in the cheap clusters, derive a relationship from the number of points to the radius $r$ of that point set, and then draw balls of radius $r$ around the points in each of those sets.
+In order to cluster the expensive points, we then use brute force to guess each of the remaining cluster centers, and since the number of remaining clusters is bounded and small, this brute force operation should not contribute significantly to the runtime.
+This all hinges on being able to partition the points into cheap and expensive sets, which arises naturally out of the procedure.
+In practice, the algorithm attempts to create a list of possible subsets of points (which could overlap in membership), and iteratively refine that list to create our $k$ clusters.
+
+We provide the algorithm in fig (TODO), and give a very high level overview of the two main "stages" included.
+
+First, the "Population stage."  We assume that we have some initial list $Q_{init}$ of graphs $\{T_i\}$ which is just the centers of the expensive clusters (i.e.
+one point per $T_i$).
+Let $n$ be the number of datapoints.
+We will count backwards (counter $s$) from $s=n$ down to $1$, i.e.
+once for each datapoint.
+
+Now, Let $d(x, Q)$ be the distance from point $x$ to any point in any of the components in $Q$.
+Now, let $r = \frac{\beta OPT}{4 s}$, where $r$ is the aforementioned ball radii; in other words, we're attempting to slowly grow the ball sizes with each iteration.
+If a point is within $2r$ of $Q$, we remove it.
+Then, we loop again over each remaining point.
+
+For each point $x$, look only at the set of points within $r$ of that point (including $x$ itself), and call this $B$.
+Then, for each pair $(a, b)$ of points in $B$, we connect them if their distance is bounded by $r$ and if there are at least $\frac{s}{2}$ points in $B$ distance $r$ from both $a$ and $b$.
+
+Finally, for each connected component that results this way, add it to $Q$ if it has at least $\frac{s}{2}$ points, and remove all the points of distance $2r$ from this component.
+This ends the population stage.
+
+TODO: What happens to the "removed" points?
+
+At the end of the population stage, we have a collection of connected components in $Q$ from which we derive clusters, in the "Centers-Retrieving stage."
+
+In this stage, we continually select candidates of $k$ components from $Q$.
+For each component, we choose the point that is the best possible center $c_i$ for that component and all points within $2r$ of that component.
+Then, we assign all points to the nearest cluster center $c_i$.
+Eventually, one of these subsets of components should give us a clustering with value at most $(1 + \epsilon) OPT$ (assuming $OPT$ is known; we will prove this).
+
+\subsection{Proof of Correctness}
 
@@
 
+\subsection{Runtime Analysis}
+
 \end{document}

### 2015-05-11 paper.tex
Through lemma 4.2

--- 2015-05-10 paper.tex
@@
 \newcommand{\vecsf}[1]{\textsf{\textbf{#1}}}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
 
@@
 \subsubsection{Bounds on Components of Cheap Clusters}
-We now move into the proof of correctness, and start by proving a useful lemma here.
-As mentioned before, our goal is to find lists of \emph{components}, or subsets of points, from which we hope we can derive our clusters.
-We now show that, if we bound
+We now move into the proof of correctness, and the paper starts by proving a useful lemma for proving the correctness of the population stage.
+In this section, we only consider operations on clusters $C_i$ which are cheap at early steps such that $s \geq |C_i|$.
+We prove two useful facts.
+
+\begin{lemma}
+Any component $T$ added in step $s$ does not contain any point $z$ such that $\frac{\beta OPT}{2 C_i} \leq  d(c_i, z) \leq \frac{3 \beta OPT}{4 C_i}$.
+\label{lemma:4.2a}
+\end{lemma}
+
+\begin{proof}
+By contradiction - assume that the distance of $z$ does fall within this range.
+In the algorithm, the $r = \frac{\beta OPT}{4 s}$, and because of the bound we imposed on $s$, $r = \frac{\beta OPT}{4 |C_i|}$.
+Let $p$ be a point within $r$ of $z$.
+If $p$ is part of $C_i$ but falls outside of the inner ring (we'll show these two facts soon), and $T$ is added to $Q$, then by step d), it must be that $T$, which is in the ball $r$ of $z$ by step c), has more than $\frac{s}{2}$, and therefore more than $\frac{C_i}{2}$ points.
+But, this is a contradiction, since more than half (at least $1 - \frac{\epsilon}{3}$) the points of any inner ring must fall \emph{within} the inner ring, not outside the inner ring, as it would here.
+
+Now, we show that $p$ is part of $C_i$ and falls outside the inner ring, completing the proof.
+
+First, $p$ is part of $C_i$.
+Since $d(c_i, z) \geq \frac{\beta OPT}{2 |C_i|}$ and $d(z, p) \leq r \leq \frac{\beta OPT}{4 |C_i|}$, then by the triangle inequality, $d(c_i, p) \geq d(c_i, z) - d(z, p) \geq \frac{\beta OPT}{4 |C_i|}$, which means that $p$ belongs to $C_i$ since our clustering is $\beta$-distributed.
+
+Second, $p$ falls outside the inner ring.
+Since $d(c_i, z) \leq \frac{3 \beta OPT}{4 |C_i|}$ and, still, $d(z, p) \leq r \leq \frac{\beta OPT}{4 |C_i|}$, by the triangle inequality, $d(c_i, p) \leq d(c_i, z) + d(z, p) \leq \frac{\beta OPT}{|C_i|}$.
+Thus, by the definition of an outer ring, $p$ must fall \emph{outside} it.
+
+\end{proof}
+
+\begin{lemma}
+Given the same preconditions, $T$ cannot contain a pair of points $(p_1, p_2)$ such that $p_1$ is within $\frac{\beta OPT}{2 |C_i|}$ of $c_i$ and $p_2$ is greater than $\frac{4 \beta OPT}{3 |C_i|}$.
+\end{lemma}
+
+\begin{proof}
+Proof by contradiction.
+Assume this did not hold.
+Since we assume that $T$ is a connected component, there's a path from $p_1$ to $p_2$.
+By step $c$ in the algorithm, two points are only connected if their distance is at most $r = \frac{\beta OPT}{4s} \leq \frac{\beta OPT}{4|C_i|}$ (the inequality following from our bound assumption on $s$ and $|C_i|$).
+Then, since we assumed this lemma did not hold, $p_1$ is within $\frac{\beta OPT}{2 |C_i|}$ of $c_i$ and $p_2$ is greater than $\frac{4 \beta OPT}{3 |C_i|}$.
+This means there can be a point that falls within that range along the connected path, which would contradict lemma \ref{lemma:4.2a}.
+\end{proof}
+
+Combined, these two lemmas state that every point in a component either lies relatively far (more than $\frac{4 \beta OPT}{3 |C_i|}$ )to $c_i$ or very close to $c_i$ (within $\frac{\beta OPT}{2 |C_i|}$).
+TODO: is this accurate?
 

### 2015-05-11 paper.tex
removing cluster size restriction

--- 2015-05-11 paper.tex
@@
 
+\subsection{Removing the cluster size restriction}
+
+Our analysis up until now has hinged on having all clusters of size at least $\delta n$.
+When lifting this restriction, there exists an algorithm which is instead $O(\delta/\alpha)$-close to the target, presented in \cite{firstpaper}.
+Because this proof is fundamentally different than those above, we will only sketch the proof for intuition.
+
+We define $w(x)$ to be the distance from a point $x$ to it's closest cluster center, and $w_2(x)$ to be the distance to the second closest center.
+Let $w = \frac{\text{OPT}}{n}$ and $d_crit = 2\frac{\alpha w}{10 \epsilon}$.
+Let $X_i$ be the set of all points in $C_i^*$ for which $w(x) < d_crit$ and $w_2(x) - w(x) \geq 5d_crit$.
+Intuitively the points in $X_i$ are those which are strongly in $C_i^*$, and so moving them to the second closest cluster is expensive.
+
+A \textbf{threshold graph} $G_\tau = (S, E_\tau)$ to be that which has all pairs of points $(x, y)$ connected if $d(x, y) \leq \tau$.
+For a problem satisfying the $(1 + \alpha, \epsilon)$-property, set $\tau = 2d_{crit}$.
+This graph has the following properties:
+
+\begin{itemize}
+\item For all $x, y \in X_i$, $(x,y) \in E(G_\tau)$. That is, if two points are both much closer to one center than any other, they will have an edge in this graph.
+
+\item If $x \in X_i$ and $y \in X_{j \neq i}$, the $(x,y) \notin E(G_\tau)$, and they share no neighbors.
+That is, if two points are tightly bound each to different clusters, they will not share edges in the threshold graph.
+
+The algorithm, taken directly from \cite{firstpaper} is: *algorithm, I'll format it like you do*
+
+As $w$ may not be known directly, running this algorithm on a $\gamma$-approximation of $w$ gives a $O(\gamma\delta/\alpha)$ approximation of the optimal clustering.
+As noted above, for sake of brevity we exclude the proof for this algorithm.
+However intuitively the threshold graph groups points into cliques,.
+By virtue of how the graph was created, we guarantee that when a cluster is selected from the graph, only a bounded number of possibly incorrect points can be added to that cluster.
+Additionally total number of points that were in a clique which belonged to a different cluster is bounded, so the final cluster also has bounded error.
+
+\end{itemize}
+
 \end{document}

### 2015-05-11 paper.tex
Through 4.3

--- 2015-05-11 paper.tex
@@
 
-\subsubsection{Bounds on Components of Cheap Clusters}
+\subsubsection{Points in components are either close or far to their centers}
 We now move into the proof of correctness, and the paper starts by proving a useful lemma for proving the correctness of the population stage.
@@
 Given the same preconditions, $T$ cannot contain a pair of points $(p_1, p_2)$ such that $p_1$ is within $\frac{\beta OPT}{2 |C_i|}$ of $c_i$ and $p_2$ is greater than $\frac{4 \beta OPT}{3 |C_i|}$.
+\label{lemma:4.2b}
 \end{lemma}
@@
 
+\subsubsection{Points in inner rings are quickly added for each cluster}
+
+Here we show that for each cluster $C_i$, $Q$ will always contain a component $T'$ with a point from the inner ring of $C_i$ by step $s = |C_i|$.
+
+To see why this is the case, assume that we are at step $s$ and the algorithm has not yet inserted some $T'$ yet.
+There are two possible things that could happen here.
+Either that component might be added at step $s$ (in which case we're done), or it could be that the point the algorithm would add has been removed by some other step.
+What we need to prove is that even if an inner ring point is removed, the algorithm will still add a different inner ring point at this step.
+
+Assume for contradiction's sake that no other inner ring point for $C_i$ is added at or before step $s$, and let $s' \geq |C_i|$ (remember we are counting \emph{backwards}) be the step at which the original inner ring point (call it $x$) were removed.
+Then if this point were removed it's due to step b), and it was removed because there already exists a point (call it $y$) in a component in $Q$ within $2r = \frac{\beta OPT}{2s'} \leq \frac{\beta OPT}{2|C_i|}$ of this $x$.
+Then, by the triangle inequality, since we also know that $d(c_i, x) \leq \frac{\beta OPT}{8 |C_i|}$ since it's an inner ring point,  $d(c_i, y) \leq d(c_i, x) + d(x, y) \leq \frac{\beta OPT}{2s} + \frac{\beta OPT}{8 |C_i|} \leq \frac{5 \beta OPT}{8 |C_i|}$.
+Now, let $s''$ be the step at which $T''$, the component which held $y$ was inserted.
+We know that $d(c_i, y) < \frac{\beta OPT}{2 |C_i|}$ (TODO: why?
+I don't understand that).
+By step d) in the algorithm, $T''$ must be of at least size $\frac{s''}{2}$ and thus at least size $\frac{|C_i|}{2}$, but by our assumption (for contradiction) does not contain a single inner ring point on $C_i$.
+Then, since we know at least half of a cheap cluster should belong to its inner ring, all of the points in $T''$ can't be in $C_i$ and so at least one point (say, $w$) must be in some $C_j~i \neq j$.
+By $\beta$ distribution, $d(c_i, w) > \frac{\beta OPT}{|C_i|}$, but by \ref{lemma:4.2b}, we can't have this far $w$ and close $y$ both in $T''$.
+Thus we have a contradiction, and so a different inner ring point than $x$ must have already been added.
+
 \subsection{Runtime Analysis}

### 2015-05-11 paper.tex

--- 2015-05-11 paper.tex
@@
 TODO: is this accurate?
+
+\subsubsection{Points in inner rings are quickly added for each cluster}
+
+Here we show that for each cluster $C_i$, $Q$ will always contain a component $T'$ with a point from the inner ring of $C_i$ by step $s = |C_i|$.
+
+To see why this is the case, assume that we are at step $s$ and the algorithm has not yet inserted some $T'$ yet.
+There are two possible things that could happen here.
+Either that component might be added at step $s$ (in which case we're done), or it could be that the point the algorithm would add has been removed by some other step.
+What we need to prove is that even if an inner ring point is removed, the algorithm will still add a different inner ring point at this step.
+
+Assume for contradiction's sake that no other inner ring point for $C_i$ is added at or before step $s$, and let $s' \geq |C_i|$ (remember we are counting \emph{backwards}) be the step at which the original inner ring point (call it $x$) were removed.
+Then if this point were removed it's due to step b), and it was removed because there already exists a point (call it $y$) in a component in $Q$ within $2r = \frac{\beta OPT}{2s'} \leq \frac{\beta OPT}{2|C_i|}$ of this $x$.
+Then, by the triangle inequality, since we also know that $d(c_i, x) \leq \frac{\beta OPT}{8 |C_i|}$ since it's an inner ring point,  $d(c_i, y) \leq d(c_i, x) + d(x, y) \leq \frac{\beta OPT}{2s} + \frac{\beta OPT}{8 |C_i|} \leq \frac{5 \beta OPT}{8 |C_i|}$.
+Now, let $s''$ be the step at which $T''$, the component which held $y$ was inserted.
+We know that $d(c_i, y) < \frac{\beta OPT}{2 |C_i|}$ (TODO: why?
+I don't understand that).
+By step d) in the algorithm, $T''$ must be of at least size $\frac{s''}{2}$ and thus at least size $\frac{|C_i|}{2}$, but by our assumption (for contradiction) does not contain a single inner ring point on $C_i$.
+Then, since we know at least half of a cheap cluster should belong to its inner ring, all of the points in $T''$ can't be in $C_i$ and so at least one point (say, $w$) must be in some $C_j~i \neq j$.
+By $\beta$ distribution, $d(c_i, w) > \frac{\beta OPT}{|C_i|}$, but by \ref{lemma:4.2b}, we can't have this far $w$ and close $y$ both in $T''$.
+Thus we have a contradiction, and so a different inner ring point than $x$ must have already been added.
 

### 2015-05-11 paper.tex
fixed up some of the proofs. Left some commented blocks for info.

@@
 
-\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size
+\documentclass[paper=a4, fontsize=10.5pt]{scrartcl} % A4 paper and 11pt font size
 
@@
 
-Define a \emph{cheap cluster} as one where $OPT_i \leq \frac{\beta \epsilon OPT}{32}$, meaning it's less than $\frac{32}{\beta \epsilon}$ of $OPT$.
+Define a \emph{cheap cluster} $C_i$ as one where $OPT_i \leq \frac{\beta \epsilon OPT}{32}$.
 Let the \emph{inner ring} of $C_i$ be the set of points $x$ such that $d(x, c_i) \leq \frac{ \beta OPT }{8 |C_i|}$.
@@
 By contradiction, assume that more than $\frac{\epsilon}{4}$ fraction of points exist outside of the inner ring.
-$OPT_i > (points outside inner ring) * (cost of points)$, and the first term in that product is lower bounded by $\frac{\epsilon |C_i|}{4}$ and that second term is lower bounded by $\frac{\beta OPT}{8 |C_i|}$.
+$OPT_i > (\text{points outside inner ring}) * (\text{cost of points})$, and the first term in that product is lower bounded by $\frac{\epsilon |C_i|}{4}$ and that second term is lower bounded by $\frac{\beta OPT}{8 |C_i|}$.
 Thus, $OPT_i > \frac{\beta \epsilon OPT }{32}$.
@@
 
-TODO: what on Earth does Markov's inequality have to do with this?  Is this a typo in the paper?
+%TODO: what on Earth does Markov's inequality have to do with this?  Is this a typo in the paper?
+
+%Yeah i have no idea... what you said seems uhh... pretty straightfoward
 
 \subsection{Algorithm}
-First, like in the paper, we'll begin by briefly discussing the intuition of the $(1 + \epsilon)$ approximation algorithm.
+We'll begin by briefly discussing the intuition of the $(1 + \epsilon)$ approximation algorithm.
 
 Let's begin by partitioning our set of clusters into two types.
-There are  "cheap" clusters and "expensive" clusters, with the idea being that the expensive clusters contribute a relatively much larger amount than the cheaper clusters; in other words, we have mostly densely packed groups of points, and a few not-so densely packed groups.
-The key idea is to assume (and later we prove that the assumption is valid) that the number of expensive clusters is \emph{bounded}.
-We then have a method for clustering the cheap point sets, assuming we can partition cheap and expensive points - guess the number of points in the cheap clusters, derive a relationship from the number of points to the radius $r$ of that point set, and then draw balls of radius $r$ around the points in each of those sets.
+There are  "cheap" clusters and "expensive" clusters, with the idea being that the expensive clusters contribute a larger amount to the cost than the cheaper clusters; in other words, we have mostly densely packed groups of points, and a few not-so densely packed groups.
+The key idea is to assume (and later we prove) that the number of expensive clusters is \emph{bounded}.
+We then have a method for clustering the cheap point sets, assuming we can partition cheap and expensive points: guess the number of points in the cheap clusters, derive a relationship from the number of points to the radius $r$ of that point set, and then draw balls of radius $r$ around the points in each of those sets.
 In order to cluster the expensive points, we then use brute force to guess each of the remaining cluster centers, and since the number of remaining clusters is bounded and small, this brute force operation should not contribute significantly to the runtime.
@@
 
-Now, Let $d(x, Q)$ be the distance from point $x$ to any point in any of the components in $Q$.
+Now, let $d(x, Q)$ be the minimum distance from point $x$ to any point in any of the components in $Q$.
 Now, let $r = \frac{\beta OPT}{4 s}$, where $r$ is the aforementioned ball radii; in other words, we're attempting to slowly grow the ball sizes with each iteration.
@@
 
-TODO: What happens to the "removed" points?
+%TODO: What happens to the "removed" points?
+%Nothing, they just aren't possible candidates for cluster centers.
 
@@
 \subsubsection{Points in components are either close or far to their centers}
-We now move into the proof of correctness, and the paper starts by proving a useful lemma for proving the correctness of the population stage.
+We now move into the proof of correctness, and the paper starts by proving a useful lemma for the correctness of the population stage.
 In this section, we only consider operations on clusters $C_i$ which are cheap at early steps such that $s \geq |C_i|$.
@@
 By contradiction - assume that the distance of $z$ does fall within this range.
-In the algorithm, the $r = \frac{\beta OPT}{4 s}$, and because of the bound we imposed on $s$, $r = \frac{\beta OPT}{4 |C_i|}$.
-Let $p$ be a point within $r$ of $z$.
+In the algorithm, $r = \frac{\beta OPT}{4 s}$, and because of the bound we imposed on $s$, $r = \frac{\beta OPT}{4 |C_i|}$.
+Let $p$ be a point such that $d(p, z) < r$.
 If $p$ is part of $C_i$ but falls outside of the inner ring (we'll show these two facts soon), and $T$ is added to $Q$, then by step d), it must be that $T$, which is in the ball $r$ of $z$ by step c), has more than $\frac{s}{2}$, and therefore more than $\frac{C_i}{2}$ points.
@@
 First, $p$ is part of $C_i$.
-Since $d(c_i, z) \geq \frac{\beta OPT}{2 |C_i|}$ and $d(z, p) \leq r \leq \frac{\beta OPT}{4 |C_i|}$, then by the triangle inequality, $d(c_i, p) \geq d(c_i, z) - d(z, p) \geq \frac{\beta OPT}{4 |C_i|}$, which means that $p$ belongs to $C_i$ since our clustering is $\beta$-distributed.
+Since $d(c_i, z) \leq \frac{3 \beta OPT}{4 |C_i|}$ and, still, $d(z, p) < r \leq \frac{\beta OPT}{4 |C_i|}$, by the triangle inequality, $d(c_i, p) \leq d(c_i, z) + d(z, p) < \frac{\beta OPT}{|C_i|}$, which means that $p$ belongs to $C_i$ since our clustering is $\beta$-distributed.
 
 Second, $p$ falls outside the inner ring.
-Since $d(c_i, z) \leq \frac{3 \beta OPT}{4 |C_i|}$ and, still, $d(z, p) \leq r \leq \frac{\beta OPT}{4 |C_i|}$, by the triangle inequality, $d(c_i, p) \leq d(c_i, z) + d(z, p) \leq \frac{\beta OPT}{|C_i|}$.
+Since $d(c_i, z) \geq \frac{\beta OPT}{2 |C_i|}$ and $d(z, p) < r \leq \frac{\beta OPT}{4 |C_i|}$, then by the triangle inequality, $d(c_i, p) \geq d(c_i, z) - d(z, p) > \frac{\beta OPT}{4 |C_i|}$.
 Thus, by the definition of an outer ring, $p$ must fall \emph{outside} it.
@@
 Since we assume that $T$ is a connected component, there's a path from $p_1$ to $p_2$.
-By step $c$ in the algorithm, two points are only connected if their distance is at most $r = \frac{\beta OPT}{4s} \leq \frac{\beta OPT}{4|C_i|}$ (the inequality following from our bound assumption on $s$ and $|C_i|$).
-Then, since we assumed this lemma did not hold, $p_1$ is within $\frac{\beta OPT}{2 |C_i|}$ of $c_i$ and $p_2$ is greater than $\frac{4 \beta OPT}{3 |C_i|}$.
+By step $c$ in the algorithm, two points are only connected if their distance is at most $r = \frac{\beta OPT}{4s} \leq \frac{\beta OPT}{4|C_i|}$.
+
+However by triangle inequality $d(p_1, p_2) \geq d(p_2, c_i) - d(p_1, c_i) \geq \frac{5\beta OPT}{6|C_i|} \geq \frac{\beta OPT}{4|C_i|}$, leading to a contradiction.
+
+%  Then, since we assumed this lemma did not hold, $p_1$ is within $\frac{\beta OPT}{2 |C_i|}$ of $c_i$ and $p_2$ is greater than $\frac{4 \beta OPT}{3 |C_i|}$.
 This means there can be a point that falls within that range along the connected path, which would contradict lemma \ref{lemma:4.2a}.
@@
 Combined, these two lemmas state that every point in a component either lies relatively far (more than $\frac{4 \beta OPT}{3 |C_i|}$ )to $c_i$ or very close to $c_i$ (within $\frac{\beta OPT}{2 |C_i|}$).
-TODO: is this accurate?
+
+%TODO: is this accurate?
+%Lemma 2.1 had a couple of clear bugs (proofs for inner ring and part of C_i were reversed) Reread 2.2 and see if I what I said made sense?
 
@@
 Now, let $s''$ be the step at which $T''$, the component which held $y$ was inserted.
-We know that $d(c_i, y) < \frac{\beta OPT}{2 |C_i|}$ (TODO: why?
-I don't understand that).
-By step d) in the algorithm, $T''$ must be of at least size $\frac{s''}{2}$ and thus at least size $\frac{|C_i|}{2}$, but by our assumption (for contradiction) does not contain a single inner ring point on $C_i$.
+We know that $d(c_i, y) < \frac{\beta OPT}{2 |C_i|}$   By step d) in the algorithm, $T''$ must be of at least size $\frac{s''}{2}$ and thus at least size $\frac{|C_i|}{2}$, but by our assumption (for contradiction) does not contain a single inner ring point on $C_i$.
 Then, since we know at least half of a cheap cluster should belong to its inner ring, all of the points in $T''$ can't be in $C_i$ and so at least one point (say, $w$) must be in some $C_j~i \neq j$.
@@
 Thus we have a contradiction, and so a different inner ring point than $x$ must have already been added.
+
+%(TODO: why?  I don't understand that).
+% if it's within frac{\beta OPT}{2|C_i|} of some x, and that x is within frac{\beta OPT}{8|C_i|}  of the center, by triangle inequality the thing you were quesitoning holds.
 

### 2015-05-13 paper.tex
Some cleanup

@@
 Here we review work done on approximations for clustering algorithms, focusing on the work of \cite{firstpaper} and \cite{secondpaper}.
-The essential goal is that various data points about objects of interest, we can group those data points by which object they belong to.
-For example we may have a collection of face images, collect relative low dimensional data about the images, then cluster this data to try to determine images came from the same people by clustering them.
-However the problem of clustering points given a distance metric in general NP-hard.
-Additionally analysis of clustering alone does not take into account the relationship between the clusters and the underlying object they belong to.
-For example even if the images were clustered optimally, that clustering may not correctly group the images by which person's face is in the image.
-The work here will focus on approximating the underlying object a data point belongs to, using clustering methods.
+The essential goal is that for various data points about objects of interest, we can group those data points by which object they belong to.
+For example we may have a collection of face images, collect a relatively low dimensional datapoint (feature vector) for each image, and then cluster this data to try to determine whether images came from the same people (each cluster corresponding to a distinct person).
+However the problem of clustering points given a distance metric is in general NP-hard.
+Additionally, analysis of clustering alone does not take into account the relationship between the clusters and the underlying object they belong to.
+For example, even if the images were clustered optimally, according to some objective function, that clustering may not correctly group the images by which person's face is in the image if the choice of objective function were not a perfect match for the problem.
+The work here will focus on algorithms for approximating correct clustering.
 
@@
 
-In order to begin our analysis of clustering, we first define what clustering means exactly, and the relationship it holds to solving an underlying data analysis problem.
+In order to begin our analysis of clustering, we first define what clustering means exactly, and the relationship it holds in solving an underlying data analysis problem.
 
@@
 \subsection{($c, \delta$)-property}
-Thus far we have only addressed clustering based on the data points, but not the underlying objectives interest.
+Thus far we have only addressed clustering based on the data points, but not the underlying objectives interest. TODO: Andy note: that's not true, we just talked about the objective above...
+
 The \textbf{objective} function $\Phi(C)$ is the cost of a clustering based only on the distance between points.
@@
 If a clustering satisfies the $(c, \delta)$-property, then for all clusterings where $\Phi(C) \leq c*\text{OPT}_\Phi$, $C$ is also $\delta$-close to $C^T$.
-By analyzing problems instances where this holds true for some values of $c$ and $\delta$, we can generate approximate clustering schema to achieve approximation solutions to discovering the target clustering.
+By analyzing problem instances where this holds true for some values of $c$ and $\delta$, we can generate approximate clustering schema to achieve approximation solutions to discovering the target clustering.
 
@@
 
-The goal of this work is to develop a minimal runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
+The goal of this work is to develop a polynomial runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
 It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{thethingthepapercitesJMS02}.
@@
 
-\item Assume we have a problem with a $(1+\alpha, \delta)$-property, and that all target clusters have size greater than $\delta n$. There exists a PTAS $\delta$-close to the target.
+\item Assume we have a problem with a $(1+\alpha, \delta)$-property, and that all target clusters have size greater than $\delta n$.
+There exists a PTAS $\delta$-close to the target and an algorithm for finding this clustering.
 
@@
 The paper begins by describing a few statistics on the point distributions which are necessary for the subsequent analysis.
-Let $OPT_i$ be the contribution to the objective function of the optimal clustering by cluster $i$, and let $OPT^{i \rightarrow j}$ denote the the cost when cluster $i$ is removed from the $OPT$ solution (TODO: unique?) and its pointed are then given to cluster $j$.
+Let $OPT_i$ be the contribution to the objective function of the optimal clustering by cluster $i$, and let $OPT^{i \rightarrow j}$ denote the the cost when cluster $i$ is removed from the $OPT$ solution (which, for simplicity's sake, we will assume to be unique) and its points are then given to cluster $j$.
 Intuitively, $OPT^{i \rightarrow j}$ corresponds to a (bad) solution of a $k-1$ clustering, and $OPT(k)^{i \rightarrow j} \geq OPT(k-1)$, where $OPT(f)$ is the optimal solution of the $f$ cluster problem.
@@
 
-Define a \emph{cheap cluster} $C_i$ as one where $OPT_i \leq \frac{\beta \epsilon OPT}{32}$.
-Let the \emph{inner ring} of $C_i$ be the set of points $x$ such that $d(x, c_i) \leq \frac{ \beta OPT }{8 |C_i|}$.
+Define a \emph{cheap cluster} $C^*_i$ as one where $OPT_i \leq \frac{\beta \epsilon OPT}{32}$.
+Let the \emph{inner ring} of $C_i$ be the set of points $x$ such that $d(x, c_i) \leq \frac{ \beta OPT }{8 |C^*_i|}$.
 If $C_i$ is cheap, then no more than $\frac{\epsilon}{4}$ points reside outside its inner ring.
 By contradiction, assume that more than $\frac{\epsilon}{4}$ fraction of points exist outside of the inner ring.
-$OPT_i > (\text{points outside inner ring}) * (\text{cost of points})$, and the first term in that product is lower bounded by $\frac{\epsilon |C_i|}{4}$ and that second term is lower bounded by $\frac{\beta OPT}{8 |C_i|}$.
+$OPT_i > (\text{points outside inner ring}) * (\text{cost of points})$, and the first term in that product is lower bounded by $\frac{\epsilon |C^*_i|}{4}$ and that second term is lower bounded by $\frac{\beta OPT}{8 |C^*_i|}$.
 Thus, $OPT_i > \frac{\beta \epsilon OPT }{32}$.
@@
 We now move into the proof of correctness, and the paper starts by proving a useful lemma for the correctness of the population stage.
-In this section, we only consider operations on clusters $C_i$ which are cheap at early steps such that $s \geq |C_i|$.
+In this section, we only consider operations on clusters $C^*_i$ which are cheap at early steps such that $s \geq |C^*_i|$.
 We prove two useful facts.
@@
 \begin{lemma}
-Any component $T$ added in step $s$ does not contain any point $z$ such that $\frac{\beta OPT}{2 C_i} \leq  d(c_i, z) \leq \frac{3 \beta OPT}{4 C_i}$.
+Any component $T$ added in step $s$ does not contain any point $z$ such that $\frac{\beta OPT}{2 C^*_i} \leq  d(c_i, z) \leq \frac{3 \beta OPT}{4 C^*_i}$.
 \label{lemma:4.2a}
@@
 By contradiction - assume that the distance of $z$ does fall within this range.
-In the algorithm, $r = \frac{\beta OPT}{4 s}$, and because of the bound we imposed on $s$, $r = \frac{\beta OPT}{4 |C_i|}$.
+In the algorithm, $r = \frac{\beta OPT}{4 s}$, and because of the bound we imposed on $s$, $r = \frac{\beta OPT}{4 |C^*_i|}$.
 Let $p$ be a point such that $d(p, z) < r$.
-If $p$ is part of $C_i$ but falls outside of the inner ring (we'll show these two facts soon), and $T$ is added to $Q$, then by step d), it must be that $T$, which is in the ball $r$ of $z$ by step c), has more than $\frac{s}{2}$, and therefore more than $\frac{C_i}{2}$ points.
+If $p$ is part of $C^*_i$ but falls outside of the inner ring (we'll show these two facts soon), and $T$ is added to $Q$, then by step d), it must be that $T$, which is in the ball $r$ of $z$ by step c), has more than $\frac{s}{2}$, and therefore more than $\frac{C_i}{2}$ points.
 But, this is a contradiction, since more than half (at least $1 - \frac{\epsilon}{3}$) the points of any inner ring must fall \emph{within} the inner ring, not outside the inner ring, as it would here.
 
-Now, we show that $p$ is part of $C_i$ and falls outside the inner ring, completing the proof.
+Now, we show that $p$ is part of $C^*_i$ and falls outside the inner ring, completing the proof.
 
-First, $p$ is part of $C_i$.
-Since $d(c_i, z) \leq \frac{3 \beta OPT}{4 |C_i|}$ and, still, $d(z, p) < r \leq \frac{\beta OPT}{4 |C_i|}$, by the triangle inequality, $d(c_i, p) \leq d(c_i, z) + d(z, p) < \frac{\beta OPT}{|C_i|}$, which means that $p$ belongs to $C_i$ since our clustering is $\beta$-distributed.
+First, $p$ is part of $C^*_i$.
+Since $d(c_i, z) \leq \frac{3 \beta OPT}{4 |C^*_i|}$ and, still, $d(z, p) < r \leq \frac{\beta OPT}{4 |C^*_i|}$, by the triangle inequality, $d(c_i, p) \leq d(c_i, z) + d(z, p) < \frac{\beta OPT}{|C^*_i|}$, which means that $p$ belongs to $C^*_i$ since our clustering is $\beta$-distributed.
 
 Second, $p$ falls outside the inner ring.
-Since $d(c_i, z) \geq \frac{\beta OPT}{2 |C_i|}$ and $d(z, p) < r \leq \frac{\beta OPT}{4 |C_i|}$, then by the triangle inequality, $d(c_i, p) \geq d(c_i, z) - d(z, p) > \frac{\beta OPT}{4 |C_i|}$.
+Since $d(c_i, z) \geq \frac{\beta OPT}{2 |C^*_i|}$ and $d(z, p) < r \leq \frac{\beta OPT}{4 |C^*_i|}$, then by the triangle inequality, $d(c_i, p) \geq d(c_i, z) - d(z, p) > \frac{\beta OPT}{4 |C^*_i|}$.
 Thus, by the definition of an outer ring, $p$ must fall \emph{outside} it.
@@
 \begin{lemma}
-Given the same preconditions, $T$ cannot contain a pair of points $(p_1, p_2)$ such that $p_1$ is within $\frac{\beta OPT}{2 |C_i|}$ of $c_i$ and $p_2$ is greater than $\frac{4 \beta OPT}{3 |C_i|}$.
+Given the same preconditions, $T$ cannot contain a pair of points $(p_1, p_2)$ such that $p_1$ is within $\frac{\beta OPT}{2 |C^*_i|}$ of $c_i$ and $p_2$ is greater than $\frac{4 \beta OPT}{3 |C^*_i|}$.
 \label{lemma:4.2b}
@@
 Since we assume that $T$ is a connected component, there's a path from $p_1$ to $p_2$.
-By step $c$ in the algorithm, two points are only connected if their distance is at most $r = \frac{\beta OPT}{4s} \leq \frac{\beta OPT}{4|C_i|}$.
+By step $c$ in the algorithm, two points are only connected if their distance is at most $r = \frac{\beta OPT}{4s} \leq \frac{\beta OPT}{4|C^*_i|}$.
 
-However by triangle inequality $d(p_1, p_2) \geq d(p_2, c_i) - d(p_1, c_i) \geq \frac{5\beta OPT}{6|C_i|} \geq \frac{\beta OPT}{4|C_i|}$, leading to a contradiction.
+However by triangle inequality $d(p_1, p_2) \geq d(p_2, c_i) - d(p_1, c_i) \geq \frac{5\beta OPT}{6|C_i|} \geq \frac{\beta OPT}{4|C^*_i|}$, leading to a contradiction.
 
@@
 
-Combined, these two lemmas state that every point in a component either lies relatively far (more than $\frac{4 \beta OPT}{3 |C_i|}$ )to $c_i$ or very close to $c_i$ (within $\frac{\beta OPT}{2 |C_i|}$).
+Combined, these two lemmas state that every point in a component either lies relatively far (more than $\frac{4 \beta OPT}{3 |C^*_i|}$ )to $c_i$ or very close to $c_i$ (within $\frac{\beta OPT}{2 |C^*_i|}$).
 
@@
 
-Here we show that for each cluster $C_i$, $Q$ will always contain a component $T'$ with a point from the inner ring of $C_i$ by step $s = |C_i|$.
+Here we show that for each cluster $C^*_i$, $Q$ will always contain a component $T'$ with a point from the inner ring of $C^*_i$ by step $s = |C^*_i|$.
 
@@
 
-Assume for contradiction's sake that no other inner ring point for $C_i$ is added at or before step $s$, and let $s' \geq |C_i|$ (remember we are counting \emph{backwards}) be the step at which the original inner ring point (call it $x$) were removed.
-Then if this point were removed it's due to step b), and it was removed because there already exists a point (call it $y$) in a component in $Q$ within $2r = \frac{\beta OPT}{2s'} \leq \frac{\beta OPT}{2|C_i|}$ of this $x$.
-Then, by the triangle inequality, since we also know that $d(c_i, x) \leq \frac{\beta OPT}{8 |C_i|}$ since it's an inner ring point,  $d(c_i, y) \leq d(c_i, x) + d(x, y) \leq \frac{\beta OPT}{2s} + \frac{\beta OPT}{8 |C_i|} \leq \frac{5 \beta OPT}{8 |C_i|}$.
+Assume for contradiction's sake that no other inner ring point for $C^*_i$ is added at or before step $s$, and let $s' \geq |C^*_i|$ (remember we are counting \emph{backwards}) be the step at which the original inner ring point (call it $x$) were removed.
+Then if this point were removed it's due to step b), and it was removed because there already exists a point (call it $y$) in a component in $Q$ within $2r = \frac{\beta OPT}{2s'} \leq \frac{\beta OPT}{2|C^*_i|}$ of this $x$.
+Then, by the triangle inequality, since we also know that $d(c_i, x) \leq \frac{\beta OPT}{8 |C^*_i|}$ since it's an inner ring point,  $d(c_i, y) \leq d(c_i, x) + d(x, y) \leq \frac{\beta OPT}{2s} + \frac{\beta OPT}{8 |C^*_i|} \leq \frac{5 \beta OPT}{8 |C^*_i|}$.
 Now, let $s''$ be the step at which $T''$, the component which held $y$ was inserted.
-We know that $d(c_i, y) < \frac{\beta OPT}{2 |C_i|}$   By step d) in the algorithm, $T''$ must be of at least size $\frac{s''}{2}$ and thus at least size $\frac{|C_i|}{2}$, but by our assumption (for contradiction) does not contain a single inner ring point on $C_i$.
-Then, since we know at least half of a cheap cluster should belong to its inner ring, all of the points in $T''$ can't be in $C_i$ and so at least one point (say, $w$) must be in some $C_j~i \neq j$.
-By $\beta$ distribution, $d(c_i, w) > \frac{\beta OPT}{|C_i|}$, but by \ref{lemma:4.2b}, we can't have this far $w$ and close $y$ both in $T''$.
+We know that $d(c_i, y) < \frac{\beta OPT}{2 |C^*_i|}$   By step d) in the algorithm, $T''$ must be of at least size $\frac{s''}{2}$ and thus at least size $\frac{|C^*_i|}{2}$, but by our assumption (for contradiction) does not contain a single inner ring point on $C^*_i$.
+Then, since we know at least half of a cheap cluster should belong to its inner ring, all of the points in $T''$ can't be in $C^*_i$ and so at least one point (say, $w$) must be in some $C_j~i \neq j$.
+By $\beta$ distribution, $d(c_i, w) > \frac{\beta OPT}{|C^*_i|}$, but by \ref{lemma:4.2b}, we can't have this far $w$ and close $y$ both in $T''$.
 Thus we have a contradiction, and so a different inner ring point than $x$ must have already been added.

### 2015-05-13 paper.tex
Fixing typo

@@
 We know that $d(c_i, y) < \frac{\beta OPT}{2 |C^*_i|}$   By step d) in the algorithm, $T''$ must be of at least size $\frac{s''}{2}$ and thus at least size $\frac{|C^*_i|}{2}$, but by our assumption (for contradiction) does not contain a single inner ring point on $C^*_i$.
-Then, since we know at least half of a cheap cluster should belong to its inner ring, all of the points in $T''$ can't be in $C^*_i$ and so at least one point (say, $w$) must be in some $C_j~i \neq j$.
+Then, since we know at least half of a cheap cluster should belong to its inner ring, all of the points in $T''$ can't be in $C^*_i$ and so at least one point (say, $w$) must be in some $C^*_j~i \neq j$.
 By $\beta$ distribution, $d(c_i, w) > \frac{\beta OPT}{|C^*_i|}$, but by \ref{lemma:4.2b}, we can't have this far $w$ and close $y$ both in $T''$.

### 2015-05-13 paper.tex
Attempt at runtime analysis

@@
 
+TODO: dscribe how we assume OPT is known and why that's valid.
+
 \subsubsection{Points in components are either close or far to their centers}
@@
 
+Step 2) in the algorithm is run $n$ times.
+For each of those $n$ runs, we loop, potentially, over all $n$  points (2 c)), and then as we look for potentially connecting points, we loop over all $n$  points again (this happens similarly in d)), leading to $O(n^3)$ iterations.
+Step 2) overall is dominated by the $O(n^3)$ time factor.
+
+Step 3) iterates over $k^{O(\frac{1}{\beta})}$ cluster guesses, and for each of these is a simple $n^2$ procedure to find the cluster centers (iterating potentially over each point $p$, and then iterating over potentially each other point to calculate statistics to determine if $p$ is the median)  (3 b).
+Subsequently, step 3 measures the cost of each of the resulting clusters, which takes $O(nk)$ time (iterate over $k$ potential clusters.
+For each, iterate over up to $n$ points).
+Thus, step 3) takes $O(n^2 + nk)$ iterations for each guess of $k$ components.
+We need to do this $k^{O(\frac{1}{\beta})}$ times, making the runtime for step 3 $O((n^2 + nk)k^{O(\frac{1}{\beta})})$
+
+Thus, the runtime for the entire algorithm is $O((n^2 + nk)k^{O(\frac{1}{\beta})} + n^3)$.
+However, this is only for each time the algorithm is run.
+In general, the algorithm must be run once for each possible guess of the expensive clusters, using a brute-force approach.
+This requires $n^{O(\frac{1}{\beta e})}$ iterations.
+Thus, the overall runtime is $O((n^2 + nk)k^{O(\frac{1}{\beta})}n^{O(\frac{1}{\beta e})})$ which is polynomial.
+
 \subsection{Removing the cluster size restriction}

### 2015-05-14 paper.tex
Typo
more proof progress

@@
 
+\subsubsection{A given component uniquely contains the inner ring of a given $C^*_i$}
+
+Consider that $T$ contains the inner ring of some $C^*_i$ (which was established above).
+We show here that all points in $T$ are within $\frac{\beta\text{OPT}}{2|C^*_i|}$ of $c_i^*$, $T \cup B(T)$ are contained in $C_I^*$ and that no other component contains inner ring points from $C^*_i$.
+
+We know all points in $T$ are within $\frac{\beta\text{OPT}}{2|C^*_i|}$ of $c_i^*$ because at least some are in the inner ring of $C^*_i$.
+Following from \ref{lemma:4.2a} and \ref{emma:4.2b}, it is then impossible for a point to exist that is further than $\frac{\beta\text{OPT}}{2|C^*_i|}$ from $c_i^*$.
+This immediately implies that $T \subset C_i^*$ because $C_i^*$ is beta distributed.
+Additionally we have that all points in B(T) are less than $2r$ away from some point in $T$.
+Because we established $2r = \frac{\beta\text{OPT}}{4s} \leq \frac{\beta\text{OPT}}{2|C_i^*|}$, all points in B(T) then within $ \frac{\beta\text{OPT}}{|C_i^*|}$ of $c_i^*$, and therefor within $C_i^*$.
+
+Because all points in $T$ are sufficiently close to $C_i^*$ there are at most $|C_i^*|$ in $T$, and so $s < 2|C_i^*|$.
+All inner ring points are at most $\frac{\beta\text{OPT}}{4|C^*_i|} < 2r$ away from any other point, requiring all points not in $T$ be added to $B(T)$.
+Therefor no other component can contain inner ring points for $C_*^i$.
+
+\subsubsection{Less than $16/(3\beta)$ that do not contain an inner ring point}
+
+Consider a $T$ which does not contain an inner ring point.
+We will show that all points in $y \in T$ have distance greater than $\frac{3\beta\text{OPT}}{8s}$ to their optimal cluster centers $c_i^*$.
+Because each $T$ contains at least $s/2$ points, by doing so we show there cannot be more than $16/3\beta$ such $T$ (as their total contribution would then be OPT).
+
+\begin{itemize}
+
+\item Assume $y$ is from an expensive cluster $C^*_i$.
+Because $Q$ are the set of expensive clusters, we know that for the point to not be thrown out $d(c^*, y) > 2r = \frac{\beta\text{OPT}}{2s} > \frac{3\beta\text{OPT}}{8s}$.
+
+\item $C^*_i$ is a cheaper cluster and $s \geq |C^*|$.
+Because we established $d(c^*, y) < \frac{\beta \text{OPT}}{2|C^*|}$ or $d(c^*, y) > \frac{3 \beta \text{OPT}}{4|C^*|}$ We know at least one point must meet the larger property.
+Since the inner ring contains $> |C^*|/2$, but $T$ also contains at least $s/2 = |C^*_i|/2$ points, some point in $T$ must not be in $C^*_i$, and therefor have distance larger than $\frac{3 \beta \text{OPT}}{4|C^*|}$.
+\ref{lemma:4.2b} then tells us all points must meet the property.
+
+\item $C^*_i$ is a cheaper cluster and $s < |C^*|$.
+We know that all points are at least $2r$ from the center, as some such component meeting the conditions of the previous section must exist.
+$d(c^*, y) \geq d(x, y) - d(c^*, x) \geq \frac{\beta\text{OPT}}{2s} - \frac{\beta\text{OPT}}{8|C^*_i|} > \frac{3\beta\text{OPT}}{8s}$
+\end{itemize}
+
 %(TODO: why?  I don't understand that).
@@
 This requires $n^{O(\frac{1}{\beta e})}$ iterations.
-Thus, the overall runtime is $O((n^2 + nk)k^{O(\frac{1}{\beta})}n^{O(\frac{1}{\beta e})})$ which is polynomial.
+Thus, the overall runtime is $O(((n^2 + nk)k^{O(\frac{1}{\beta})}+n^3)n^{O(\frac{1}{\beta e})}) = O(n^{O(1/\beta\epsilon)}k^{O(1/\beta)}$ .
 

### 2015-05-14 paper.tex
finished proof, will clean up in a bit

@@
 We know all points in $T$ are within $\frac{\beta\text{OPT}}{2|C^*_i|}$ of $c_i^*$ because at least some are in the inner ring of $C^*_i$.
-Following from \ref{lemma:4.2a} and \ref{emma:4.2b}, it is then impossible for a point to exist that is further than $\frac{\beta\text{OPT}}{2|C^*_i|}$ from $c_i^*$.
+Following from \ref{lemma:4.2a} and \ref{lemma:4.2b}, it is then impossible for a point to exist that is further than $\frac{\beta\text{OPT}}{2|C^*_i|}$ from $c_i^*$.
 This immediately implies that $T \subset C_i^*$ because $C_i^*$ is beta distributed.
@@
 \end{itemize}
+
+\subsubsection{The algorithm outputs a k-clustering with cost $\leq (1+\epsilon)OPT$}
+
+In order to show this we will compare the total cost of clusters $\mathcal{C}$ Derived by $T_1, ..., T_k$ with those of $\mathcal{C}^*$.
+We know that the expensive clusters $C_i$ have centers $c_i^*$, and so its total cost will not exceed $\text{OPT}_i$.
+
+The cost of a cheap cluster is $\sum\limits_{x\in T\cup B(T)}d(x, c_i) + \sum\limits_{x \in C_i^* and x\notin T\cup B(T)}d(x, c_i)$.
+Because all points in the first term are in the innter ring, that includes the center, so $\sum\limits_{x\in T\cup B(T)}d(x, c_i) \leq \sum\limits_{x\in T\cup B(T)}d(x, c^*_i)$.
+We must show the error in $\sum\limits_{x \in C_i^* and x\notin T\cup B(T)}d(x, c_i)$ is bounded by $(1+\epsilon)$.
+
+We will show that $d(c_i, c_i*) < \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$ for a cheap cluster.
+We know by the definition of a cheap cluster that the cost is at most $(\epsilon/32)\beta\text{OPT}/|C_i^*|$.
+In order not exceed this bound, at least half of the points have distance at most $(\epsilon/16)\beta\text{OPT}/|C_i^*|$ from $c_i^*$.
+Also these are by definition in the inner ring.
+We know that at least half the points in a cheap cluster reside in the inner ring.
+If $d(c_i, c_i^*)  \geq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$, then we require at least half of the points to contribute at least $(\epsilon/16)\beta\text{OPT}/|C_i^*|$ to the sum, which means some point in the inner ring must do so.
+Therefor $\sum\limits_{x\in T\cup B(T)}d(x, c_i)> OPT_i$, which contradicts the optimality of the inner ring points.
+
+We therefor for each $x \in C_i^* and x\notin T\cup B(T)$ we have $d(c_i, c_i^*) \leq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$.
+So $d(x, c_i) \leq d(x, c_i^*) + d(c_i^*, c_i) \leq (1+\epsilon)d(x,c_i^*)$.
+And so each point adds at most a factor of $(1+\epsilon)$ to the error, bounding our total error to $(1+\epsilon)\text{OPT}$
 

### 2015-05-14 paper.tex
citations
some cleanup

@@
 Here we review work done on approximations for clustering algorithms, focusing on the work of \cite{firstpaper} and \cite{secondpaper}.
-The essential goal is that for various data points about objects of interest, we can group those data points by which object they belong to.
-For example we may have a collection of face images, collect a relatively low dimensional datapoint (feature vector) for each image, and then cluster this data to try to determine whether images came from the same people (each cluster corresponding to a distinct person).
+The essential goal is that for various data points about sets of interest, we wish to group those data points by which set they belong to.
+For example we may have a collection of face images, generate a relatively low dimensional datapoint (feature vector) for each image, and then cluster this data to try to determine whether images came from the same people (each cluster corresponding to a distinct person).
 However the problem of clustering points given a distance metric is in general NP-hard.
-Additionally, analysis of clustering alone does not take into account the relationship between the clusters and the underlying object they belong to.
-For example, even if the images were clustered optimally, according to some objective function, that clustering may not correctly group the images by which person's face is in the image if the choice of objective function were not a perfect match for the problem.
-The work here will focus on algorithms for approximating correct clustering.
+Additionally, analysis of clustering alone does not take into account the relationship between the clusters and the underlying sets they belong to.
+For example, even if the images were clustered optimally, according to some objective function of distance between datapoints, that clustering may not correctly group the images by which person's face is in the image.
+The work here will focus on algorithms for approximating correct clustering based on the underlying sets.
 
@@
 \subsection{($c, \delta$)-property}
-Thus far we have only addressed clustering based on the data points, but not the underlying objectives interest. TODO: Andy note: that's not true, we just talked about the objective above...
+Thus far we have only addressed clustering based on the data points, but not the underlying sets of interest.
 
 The \textbf{objective} function $\Phi(C)$ is the cost of a clustering based only on the distance between points.
-However the optimal \textbf{target} clustering is that which partitions the points best based on the underlying objects the data points are drawn from.
+However the optimal \textbf{target} clustering is that which partitions the points best based on the underlying sets the data points are drawn from.
 In the ideal problem statement the optimal objective clustering $C^*$ and the optimal target clustering $C^T$ are the same.
@@
 The goal of this work is to develop a polynomial runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
-It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{thethingthepapercitesJMS02}.
+It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{JMS02}.
 In this review we will show explore the following results:
@@
 
+\item Assume we have a problem with a $(1+\alpha, \delta)$-property, with no restriction on cluster size.
+There exists n $O(\delta/\alpha)$-close to the target and an algorithm for finding this clustering.
 \end{itemize}
@@
 This paper looks at a few statistics about how points are distributed and how it relates to optimal clusterings.
-In particular, it studies notions of "stability," which describes bounds on how much adding or removing clusters from a clustering solution changes the objective function's value, and separability, which describes how far away clusters of points are from one another.
-By characterizing point sets in terms of these statistics, the authors derive an algorithm for efficiently clustering the points and derive guarantees on the cluster quality, as well as prove its correctness.
-
+In particular, it studies notions of stability, which describes bounds on how much adding or removing clusters from a clustering solution changes the objective function's value, as well as the notion of separability, which describes how far away clusters of points are from one another.
+By characterizing point sets in terms of these statistics, the authors derive an algorithm for efficiently clustering the points and derive guarantees on the cluster quality.
 \subsection{Definitions}
@@
 Let $OPT_i$ be the contribution to the objective function of the optimal clustering by cluster $i$, and let $OPT^{i \rightarrow j}$ denote the the cost when cluster $i$ is removed from the $OPT$ solution (which, for simplicity's sake, we will assume to be unique) and its points are then given to cluster $j$.
-Intuitively, $OPT^{i \rightarrow j}$ corresponds to a (bad) solution of a $k-1$ clustering, and $OPT(k)^{i \rightarrow j} \geq OPT(k-1)$, where $OPT(f)$ is the optimal solution of the $f$ cluster problem.
+Intuitively, $OPT^{i \rightarrow j}$ corresponds to a (bad) solution of a $k-1$ clustering, and $OPT(k)^{i \rightarrow j} \geq OPT(k-1)$, where $OPT(f)$ is the optimal solution of the $f$-cluster problem.
 Let a clustering be $(1 + \alpha)$ weak-deletion stable if, $\forall i \neq j, OPT^{i \rightarrow j} > (1 + \alpha)OPT$ - in other words, removing any cluster and remapping those points to another single cluster severely contributes to the objective function, suggesting that that each cluster is significant to the solution.
@@
 
-Let's begin by partitioning our set of clusters into two types.
-There are  "cheap" clusters and "expensive" clusters, with the idea being that the expensive clusters contribute a larger amount to the cost than the cheaper clusters; in other words, we have mostly densely packed groups of points, and a few not-so densely packed groups.
+To complete the notion of a cheap cluster, we define an expensive cluster to be one with $\text{OPT}_i > \frac{\beta\epsilon\text{OPT}}{32}$.
 The key idea is to assume (and later we prove) that the number of expensive clusters is \emph{bounded}.
@@
 \end{itemize}
+\bibliography{biblio}{}
+\bibliographystyle{plain}
 

### 2015-05-14 paper.tex
Adding alg box

--- 2015-05-14 paper.tex
@@
 \usepackage{amsmath,amsfonts,amsthm} % Math packages
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 \usepackage{mathtools}
@@
 \usepackage{parskip}
+
+\makeatletter
+\def\BState{\State\hskip-\ALG@thistlm}
+\makeatother
 
@@
 \subsection{Algorithm}
+
+\begin{algorithm}
+\caption{Cluster TODO: better name}\label{alg:cluster}
+\begin{algorithmic}[1]
+\Procedure{Cluster}{}
+\BState \textbf{Initialization Stage:} $Q \gets Q_{init}$
+\BState \textbf{Population Stage:} For $s=n, n-1, n-2, \ldots 1$ do:
+\State a) Set $r=\frac{\beta OPT}{4 s}$
+\State b) Remove any point $x$ such that $d(x, Q) < 2r$.
+\State c) For any remaining data point $x$, denote the set of data points whose distance from $x$ is at most $r$ by $B(x, r)$.
+Connect any two remaining points $a$ and $b$ if: (i) $d(a, b) \lq r$ (ii) $|B(a, r)| > \frac{s}{2}$ and (iii) $|B(b, r)| > \frac{s}{2}$.
+\State d) Let $T$ be a connected component of size $ > \frac{s}{2}$.
+Then i) Add $T$ to $Q$.
+ii) Define the set $B(T) = \{x : d(x, y) \leq 2r$ for some $y \in T\}$.
+Remove the points $B(T)$ from the instance.
+\BState \textbf{Centers-Retrieving Stage}: For any choice of $k$ components $T_1, T_2, \ldots, T_k$ out of $Q$:
+\State a) Find the best center $C_i$ for $T_i \cup B(T_i)$.
+\State b) Partition all $n$ points according to the nearest point among the $k$ centers of the current $k$ components.
+\State c) If a clustering of at most $(1 + \epsilon)OPT$ is found - output these $k$ centers and halt.
+ %\hspace{\algorithmicindent} $pq.put( (v_j, e_{jk}), e_{jk}^z)$.
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
 We'll begin by briefly discussing the intuition of the $(1 + \epsilon)$ approximation algorithm.

### 2015-05-14 paper.tex
Both algs added now

@@
 
-We provide the algorithm in fig (TODO), and give a very high level overview of the two main "stages" included.
+We provide the algorithm in Alg. \ref{alg:cluster}, and give a very high level overview of the two main "stages" included.
 
@@
 
+\begin{algorithm}
+\caption{k-Mean Clustering in the General Case}\label{alg:cluster2}
+\begin{algorithmic}[1]
+\Procedure{Cluster}{}
+\BState \textbf{Input:} $w, \delta \leq 1, \alpha > 0, k.$
+\BState \textbf{Step 1:} Construct the $\tau$-threshold graph $G_{\tau}$ with $\tau = 2d_{crit} = \frac{\alpha w}{5 \delta}$.
+
+\BState \textbf{Step 2:} For $j = 1$ to $k$ do:
+\State Pick the vertex $v_j$ of the highest degree in $G_{\tau}$.
+\State Remove $v_j$ and its neighborhood from $G_{\tau}$ and call this cluster $C(v_j)$.
+\BState \textbf{Step 3:} Output the $k$ clusters $C(v_1), ..., C(v_{k-1}), S - \bigcup_{i=1}^{k-1}C(v_i)$.
+
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
 Our analysis up until now has hinged on having all clusters of size at least $\delta n$.

### 2015-05-14 paper.tex
some cleanup

--- 2015-05-14 paper.tex
@@
 Consider that $T$ contains the inner ring of some $C^*_i$ (which was established above).
-We show here that all points in $T$ are within $\frac{\beta\text{OPT}}{2|C^*_i|}$ of $c_i^*$, $T \cup B(T)$ are contained in $C_I^*$ and that no other component contains inner ring points from $C^*_i$.
+We show here that all points in $T$ are within $\frac{\beta\text{OPT}}{2|C^*_i|}$ of $c_i^*$, $T \cup B(T)$ are contained in $C_i^*$ and that no other component contains inner ring points from $C^*_i$.
 
@@
 
-Because all points in $T$ are sufficiently close to $C_i^*$ there are at most $|C_i^*|$ in $T$, and so $s < 2|C_i^*|$.
+Because all points in $T$ are sufficiently close to $C_i^*$ there are at most $|C_i^*|$ points in $T$, and so $s < 2|C_i^*|$.
 All inner ring points are at most $\frac{\beta\text{OPT}}{4|C^*_i|} < 2r$ away from any other point, requiring all points not in $T$ be added to $B(T)$.
-Therefor no other component can contain inner ring points for $C_*^i$.
+Therefor no other component can contain inner ring points for $C_i^*$.
 
@@
 In order to show this we will compare the total cost of clusters $\mathcal{C}$ Derived by $T_1, ..., T_k$ with those of $\mathcal{C}^*$.
-We know that the expensive clusters $C_i$ have centers $c_i^*$, and so its total cost will not exceed $\text{OPT}_i$.
+
+ We know that each expensive cluster $C_i$ has centers $c_i^*$, and so its total cost will not exceed $\text{OPT}_i$.
 
 The cost of a cheap cluster is $\sum\limits_{x\in T\cup B(T)}d(x, c_i) + \sum\limits_{x \in C_i^* and x\notin T\cup B(T)}d(x, c_i)$.
-Because all points in the first term are in the innter ring, that includes the center, so $\sum\limits_{x\in T\cup B(T)}d(x, c_i) \leq \sum\limits_{x\in T\cup B(T)}d(x, c^*_i)$.
+Because all points in the first term are in the inner ring, that includes the center, so $\sum\limits_{x\in T\cup B(T)}d(x, c_i) \leq \sum\limits_{x\in T\cup B(T)}d(x, c^*_i)$.
 We must show the error in $\sum\limits_{x \in C_i^* and x\notin T\cup B(T)}d(x, c_i)$ is bounded by $(1+\epsilon)$.
@@
 This requires $n^{O(\frac{1}{\beta e})}$ iterations.
-Thus, the overall runtime is $O(((n^2 + nk)k^{O(\frac{1}{\beta})}+n^3)n^{O(\frac{1}{\beta e})}) = O(n^{O(1/\beta\epsilon)}k^{O(1/\beta)}$ .
+Thus, the overall runtime is $O(((n^2 + nk)k^{O(\frac{1}{\beta})}+n^3)n^{O(\frac{1}{\beta e})}) = O(n^{O(1/\beta\epsilon)}k^{O(1/\beta))}$ .
 
@@
 Let $w = \frac{\text{OPT}}{n}$ and $d_crit = 2\frac{\alpha w}{10 \epsilon}$.
-Let $X_i$ be the set of all points in $C_i^*$ for which $w(x) < d_crit$ and $w_2(x) - w(x) \geq 5d_crit$.
+Let $X_i$ be the set of all points in $C_i^*$ for which $w(x) < d_{crit}$ and $w_2(x) - w(x) \geq 5d_{crit}$.
 Intuitively the points in $X_i$ are those which are strongly in $C_i^*$, and so moving them to the second closest cluster is expensive.

### 2015-05-14 paper.tex
uhhh yeah so we don't need opt great

@@
 Then, we assign all points to the nearest cluster center $c_i$.
-Eventually, one of these subsets of components should give us a clustering with value at most $(1 + \epsilon) OPT$ (assuming $OPT$ is known; we will prove this).
+Eventually, one of these subsets of components should give us a clustering with value at most $(1 + \epsilon) OPT$, assuming OPT is known.
+While the algorithm could be shown to work using common estimators for OPT, they will be excluded from analysis.
 
 \subsection{Proof of Correctness}
-
-TODO: dscribe how we assume OPT is known and why that's valid.
 
@@
 
-The algorithm, taken directly from \cite{firstpaper} is: *algorithm, I'll format it like you do*
+\ref{alg:cluster2} taken directly from \cite{firstpaper} uses such a threshold graph to generate the clusters.
 

### 2015-05-14 paper.tex
Minor edits

@@
 The optimal objective clustering is that which minimizes $\Phi(C) = \sum\limits_{i = 1}^k \sum_{x \in C_i} d(x, c_i)$.
+For $k$-medians, $d$ is the $L_1$ norm.
 
@@
 
-The goal of this work is to develop a polynomial runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
+The goal of these papers is to develop a polynomial runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above.
 It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{JMS02}.
@@
 
-\item Assume that the (k-1)-medians optimal cluster is more expensive than the k optimal clustering by a factor of $(1+\alpha)$ where $\alpha > 0$.
+\item Assume that the ($k-1$)-medians optimal cluster is more expensive than the $k$ optimal clustering by a factor of $(1+\alpha)$ where $\alpha > 0$.
 Then it is possible to achieve a ($1+\epsilon$) approximation to the optimal k-medians in time polynomial in $n$ and $k$, and exponential in $1/\epsilon$ and $1/\alpha$, where $n = |S|$.
@@
 \begin{algorithm}
-\caption{Cluster TODO: better name}\label{alg:cluster}
+\caption{$k$-Median Approximation Clustering (as in \cite{secondpaper})}\label{alg:cluster}
 \begin{algorithmic}[1]
@@
 Additionally we have that all points in B(T) are less than $2r$ away from some point in $T$.
-Because we established $2r = \frac{\beta\text{OPT}}{4s} \leq \frac{\beta\text{OPT}}{2|C_i^*|}$, all points in B(T) then within $ \frac{\beta\text{OPT}}{|C_i^*|}$ of $c_i^*$, and therefor within $C_i^*$.
+Because we established $2r = \frac{\beta\text{OPT}}{4s} \leq \frac{\beta\text{OPT}}{2|C_i^*|}$, all points in B(T) then within $ \frac{\beta\text{OPT}}{|C_i^*|}$ of $c_i^*$, and therefore within $C_i^*$.
 
@@
 All inner ring points are at most $\frac{\beta\text{OPT}}{4|C^*_i|} < 2r$ away from any other point, requiring all points not in $T$ be added to $B(T)$.
-Therefor no other component can contain inner ring points for $C_i^*$.
+Therefore no other component can contain inner ring points for $C_i^*$.
 
@@
 Because we established $d(c^*, y) < \frac{\beta \text{OPT}}{2|C^*|}$ or $d(c^*, y) > \frac{3 \beta \text{OPT}}{4|C^*|}$ We know at least one point must meet the larger property.
-Since the inner ring contains $> |C^*|/2$, but $T$ also contains at least $s/2 = |C^*_i|/2$ points, some point in $T$ must not be in $C^*_i$, and therefor have distance larger than $\frac{3 \beta \text{OPT}}{4|C^*|}$.
+Since the inner ring contains $> |C^*|/2$, but $T$ also contains at least $s/2 = |C^*_i|/2$ points, some point in $T$ must not be in $C^*_i$, and therefore have distance larger than $\frac{3 \beta \text{OPT}}{4|C^*|}$.
 \ref{lemma:4.2b} then tells us all points must meet the property.
@@
 If $d(c_i, c_i^*)  \geq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$, then we require at least half of the points to contribute at least $(\epsilon/16)\beta\text{OPT}/|C_i^*|$ to the sum, which means some point in the inner ring must do so.
-Therefor $\sum\limits_{x\in T\cup B(T)}d(x, c_i)> OPT_i$, which contradicts the optimality of the inner ring points.
+Therefore $\sum\limits_{x\in T\cup B(T)}d(x, c_i)> OPT_i$, which contradicts the optimality of the inner ring points.
 
-We therefor for each $x \in C_i^* and x\notin T\cup B(T)$ we have $d(c_i, c_i^*) \leq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$.
+We therefore for each $x \in C_i^* and x\notin T\cup B(T)$ we have $d(c_i, c_i^*) \leq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$.
 So $d(x, c_i) \leq d(x, c_i^*) + d(c_i^*, c_i) \leq (1+\epsilon)d(x,c_i^*)$.
